---
title: "2026_SCGL_work"
author: "Jonathan Amegin"
date: "2026-05-01"
output: html_document
editor_options: 
  markdown: 
    wrap: 90
---

Updates by Michelle DePrenger-Levin 2026.06.02

To do:

~~- for each observation, create clones in the data frame that switch the
"direction_from_transect" field for later duplicate analysis~~

~~- Split the comparison of group sizes up between seedlings, vegetative, and
reproductive~~

~~-   Develop probability scoring to show 'percent similarity' between observations based on
    weighted scores of covariates

~~-   Keep only the duplicates with the shortest distance between the compared observations
    so that only 1 duplicate/global_id

##Setting Up

```{r message=FALSE, warning=FALSE, include=FALSE}
# tidyverse contains both 'dyplr' and 'ggplot2'
library(tidyverse)
library(janitor)
library(uuid)

setwd("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Sclerocactus-glaucus/Sclerocactus-glaucus_RawData/2024_CrossedDistanceSampling")
#this contains all observations of clusters of SCGL for all sites & Transects
scgl_obs <- read_csv("distanceSampling_1.csv")
scgl_obs <- clean_names(scgl_obs)

#This contains metadata for each transect including length & direction
scgl_trans_length <- read_csv("transect_2.csv")
scgl_trans_length <- clean_names(scgl_trans_length)

#filtering NA's and outliers, and changing some variable classes for later work
scgl <- scgl_obs[!(is.na(scgl_obs$transect_meter_mark|scgl_obs$distance_to_group)),] %>%  
        filter(transect_meter_mark < 110, distance_to_group < 10) %>% 
        mutate( object_id = as.factor(object_id),
                site_name = as.factor(site_name),
                transect_number = as.factor(transect_number),
                swapped_dir = 0)

scgl[is.na(scgl[,"number_seedlings_in_group"]), "number_seedlings_in_group"] <- 0
scgl[is.na(scgl[,"number_vegetative_in_group"]), "number_vegetative_in_group"] <- 0
scgl[is.na(scgl[,"number_reproductive_in_group"]), "number_reproductive_in_group"] <- 0

#adding clone observations with the 'direction_to_group' swapped for later analysis
#some observations had NA, so need to check both directions for them
### The NA+ and NA-  see lines 74, default is + for direction
scgl1 <- rbind(scgl,
               mutate(scgl,
                      swapped_dir = "1",
                      direction_to_group = case_when(direction_to_group == "+" ~ "-",
                                                     direction_to_group == "-" ~ "+",
                                                     is.na(direction_to_group) ~ "NA-"
                                                     )))

scgl1[is.na(scgl1[,"direction_to_group"]), "direction_to_group"] <- "NA+"

#use the attributes to add modifiers to the measurements
#EW transect meter marks need to be * -1 because they were recorded as if in the(+,+) quadrant, but the mapping we will do is based on (-,+) quadrant.             
scgl1 <-
  scgl1 %>%
  mutate(
    transect_meter_mark = case_when(
      transect_orientation == "EW" ~ transect_meter_mark * -1,
      TRUE ~ transect_meter_mark),
    #making the 'direction to group' indicate if the 'distance to group' needs to be left/right of transect.  
    distance_to_group = case_when(
      direction_to_group == "-" | direction_to_group == "NA-" ~ distance_to_group * -1,
                                                         TRUE ~ distance_to_group
      ))
           
#setting up the (X,Y) columns to then transform based on plot/shapefile analysis
scgl1 <- scgl1 %>% 
  mutate( x = case_when(transect_orientation == "NS" ~ distance_to_group,
                        transect_orientation == "EW" ~ transect_meter_mark), 
          y = case_when(transect_orientation == "EW" ~ distance_to_group,
                        transect_orientation == "NS" ~ transect_meter_mark)) 

#joining the two dataframes to add transect metadata to each observation
scgl1 <- left_join(scgl1, scgl_trans_length[, c("transect_length", "global_id")],
                   by = join_by(parent_global_id == global_id))                
                
#filtering to sites
fram <- scgl1 %>% 
  filter(site_name == "12")

picnic <- scgl1 %>% 
  filter(site_name == "1"| site_name == "3")

bridgeport <- scgl1 %>% 
  filter(site_name == "10")

oilpad <- scgl1 %>% 
  filter(site_name == "11")

```



##picnic
```{r}
#transforming coordinates to match the plot design. This requires accurate note taking
 picnic_1 <-  picnic %>% 
   mutate( 
    x = case_when(transect_orientation == "NS" & transect_number == "1" ~ x - 20, 
                  transect_orientation == "NS" & transect_number == "2" ~ x - 30,
                  transect_orientation == "NS" & transect_number == "3" ~ x - 40,
                  transect_orientation == "EW" & transect_number == "1" ~ x - 20,
                  transect_orientation == "EW" & transect_number == "2" ~ x - 10,
                  transect_orientation == "EW" & transect_number == "3" ~ x - 10,
                  transect_orientation == "EW" & transect_number == "4" ~ x
                  ),
    y = case_when(transect_orientation == "NS" & transect_number == "1" ~ y + 20, 
                  transect_orientation == "NS" & transect_number == "2" ~ y + 10,
                  transect_orientation == "NS" & transect_number == "3" ~ y ,
                  transect_orientation == "EW" & transect_number == "1" ~ y + 20,
                  transect_orientation == "EW" & transect_number == "2" ~ y + 30,
                  transect_orientation == "EW" & transect_number == "3" ~ y + 40,
                  transect_orientation == "EW" & transect_number == "4" ~ y + 50,
                  ))

#adding an attribute to show if an observation is within the zone that NS & EW transects share
picnic_1 <-
  mutate(picnic_1,
         zoned = case_when(
           c(rowSums(mapply(FUN = function(a, b){between(a, b-3, b+3)}, picnic_1[, "y"], c(20, 30, 40, 50))) +
             rowSums(mapply(FUN = function(a, b){between(a, b-3, b+3)}, picnic_1[, "x"], c(-20, -30, -40)))) == 2 ~ 1,
           TRUE ~ 0))

## Add unique ID for each zone, only compare plants within a zone, not across all plants
# as a factor, fine. Where did object_id 19 come from???
picnic_1 %>%
  filter(object_id == 236)
```

### 2026 function

Idea is to use covariates to generate a probablity score indicating likelihood of pairs of
observations being duplicates.

Covariates affecting probability:

-   Number in groups, split between categories.

-   Distance between observations

-   Direction of transect

-   within/out the crossed area


```{r}
###Custom Function 1
# Input plot data mapped to (X, Y) grid and plants added with the direction_to_group swapped - test uncertainty of direction
dup.fun <- function(site, Tdistance = 1){ # , Tcluster = 12, TS = 8, TV = 2, TR = 2){
  
  site <- site %>%
    mutate_at(c("object_id"), ~as.numeric(as.character(.)))
  
  #making new column with new UUID for each observation due to cloned entries
  site$new_id = UUIDgenerate(use.time = FALSE, n = nrow(site), output = "string")
  
  #Creating list of all combinations of global_id's to reference, order doesn't matter, start with only original
  list <- utils::combn(site$new_id[site$swapped_dir == 0], m = 2, simplify = FALSE)

  # (1) first go through and keep or discard a swapped plant placement,
  ## But, doesn't matter if they're on the same transect, doesn't matter if they are far away, Only getting duplicate, use that to 
  #Appending calculated variables to the list  
  list <- lapply(list, function(list.1){

    #indexing data with the paired global_ids
    a <- site[site$new_id == list.1[1],]
    b <- site[site$new_id == list.1[2],]

    # Check the swapped 
    a1 <- site[site$object_id == a$object_id & site$swapped_dir == 1,]
    b1 <- site[site$object_id == b$object_id & site$swapped_dir == 1,]

    #Calculating distance, if the swapped is closer, then always use the swapped, only if on different transects
    # do both for all, then only keep the original or swapped if_any are marked that the swapped is closer
    ## Need a case_when for both
    #  They could be a duplicate if (1) they are from different transect orientations
    #                               (2) the pair (or any combination of swapped directions) are closer than the threshold distance 
    ### In this case "a" and "b" represent the original mapped position, "a1" and "b1" are the swapped direction
    # if(a$transect_orientation != b$transect_orientation){
      Original <- dist(rbind(c(a$x,a$y), c(b$x,b$y)), method = "euclidean")
      SwappedA <- dist(rbind(c(a1$x,a$y), c(b$x,b$y)), method = "euclidean")
      SwappedB <- dist(rbind(c(a$x,a$y), c(b1$x,b$y)), method = "euclidean")
      SwappedAB <- dist(rbind(c(a1$x,a$y), c(b1$x,b$y)), method = "euclidean")
      
      if(any(c(Original,SwappedA,SwappedB,SwappedAB) < Tdistance)){
        ## Which should be swapped
        SwapMe <- c("a_b","a1_b","a_b1","a1_b1")[which.min(c(Original,SwappedA,SwappedB,SwappedAB))]
        list.1[3] <- min(Original, SwappedA, SwappedB, SwappedAB)
      } else {
        list.1[3] <- dist(rbind(c(a$x,a$y), c(b$x,b$y)), method = "euclidean")
        SwapMe <- "a_b"
      }
      

    #determining if they are on the same transect
    ## same == 1, different == 0
    list.1[4] <- case_when(a$transect_orientation == b$transect_orientation ~ 1,
                           a$transect_orientation != b$transect_orientation ~ 0)

    
    #Comparing if the 'direction_from_transect' variable was swapped earlier
    list.1[5] <- case_when(any(sapply(SwapMe, str_detect, "a1")) ~ a1$new_id,
                           TRUE ~ a$new_id) ## id_1 updated
    list.1[6] <- case_when(any(sapply(SwapMe, str_detect, "b1")) ~ b1$new_id,
                           TRUE ~ b$new_id) ## id_2 updated
    #The orientation of id_1; id_2
    list.1[7] <- a$transect_orientation
    list.1[8] <- b$transect_orientation
    
    # First or second swapped? 
    list.1[9] <- case_when(any(sapply(SwapMe, str_detect, "a1")) ~ 1,
                           TRUE ~ 0)
    list.1[10] <- case_when(any(sapply(SwapMe, str_detect, "b1")) ~ 1,
                           TRUE ~ 0) ## id_1_updated
    
    list.1[11] <- a$object_id
    list.1[12] <- b$object_id
    
     return(list.1)
  })
  

    df <- data.frame(do.call(rbind,list))
    names(df) <- c("V1_id", "V2_id",
                      "distanceBetween",
                      "same_diff",  ## if on same or different transect 1 = same, 0 = different
                      "V1_updatedid",  # after checking if need to be swapped
                      "V2_updatedid",  # after checking if need to be swapped
                      "V1_orientation",
                      "V2_orientation",
                      "V1_swapped",
                      "V2_swapped",
                      "V1_objectID",
                      "V2_objectID")

 df$dupID <- UUIDgenerate(use.time = FALSE, n = nrow(df), output = "string") 


 ## df wide to long splitting transect orientation and id; keep only the swapped or original. 
 df_long <- df %>%
   filter(same_diff == 0) %>%
   filter(distanceBetween < Tdistance) %>%
   pivot_longer(cols = starts_with("V"), names_to = c("ID", ".value"),
                names_sep = "_") %>%
   mutate_at(c("objectID","distanceBetween","same_diff"), ~as.numeric(as.character(.))) %>%
   group_by(objectID) %>%
   filter(distanceBetween == min(distanceBetween)) %>%
   arrange(desc(swapped)) %>%
   filter(row_number() == 1) %>%
   ungroup() %>%
   select(!c(ID, same_diff)) 

 # The updated ID is the swapped or original (if swapped is < Tdistance) 
 # the dupID should be applied to the updatedid when < Tdistance
 siteout <- site %>%
   left_join(df_long, by = c("new_id" = "updatedid")) %>%
   mutate(dupID = case_when(is.na(dupID) ~ new_id,
                            TRUE ~ dupID)) %>%
   mutate(objectID = case_when(is.na(objectID) ~ object_id,
                               TRUE ~ objectID)) %>%
   group_by(object_id) %>%
   ## need to keep only swapped_dir == 0 unless swapped was true (1)
   filter(case_when(swapped == 1 ~ swapped_dir == 1,
                    TRUE ~ swapped_dir == 0)) %>%
   ungroup() %>%
   as.data.frame()
   
 return(siteout)
}

```

##Fram

###Plotting individuals

Transforming Fram coordinates to accurately reflect plot design.This is based on the
origin (0,0) being in the lower right of the plot so that one or both of the axis matches
the start of at least one transect. In this case, the origins of the three NS transects
are at Y = 0, and 4 out of 5 EW transects' origins are at X = 0.

```{r fram transform}

fram_1 <- fram %>% 
  mutate( 
    x = case_when(transect_orientation == "NS" & transect_number == "1" ~ x - 10, 
                  transect_orientation == "NS" & transect_number == "2" ~ x - 20,
                  transect_orientation == "NS" & transect_number == "3" ~ x - 30,
                  transect_orientation == "EW" & transect_number == "5" ~ x - 10,
                  TRUE ~ x
                  ),
    y = case_when(transect_orientation == "EW" & transect_number == "1" ~ y + 10,
                  transect_orientation == "EW" & transect_number == "5" ~ y + 50,
                  transect_orientation == "EW" & transect_number == "7" ~ y + 70,
                  transect_orientation == "EW" & transect_number == "8" ~ y + 80,
                  transect_orientation == "EW" & transect_number == "9" ~ y + 90,
                  TRUE ~ y),
    )

#Fram NS lines tapes were laid backwards making 0 = 100. 
fram_1 <- fram_1 %>% 
  mutate( y = case_when(transect_orientation =="NS" ~ ((y - 100)* -1), TRUE ~ y)
    )

fram_1 <- mutate(fram_1, zoned = 
                   case_when(
                     c(rowSums(mapply( FUN = function(a, b){between(a, b-3, b+3)}, fram_1[, "y"], c(10, 50, 70, 80, 90))) +
                                                rowSums(mapply( FUN = function(a, b){between(a, b-3, b+3)}, fram_1[, "x"], c(-10, -20, -30)))) == 2 ~ 1,
       TRUE ~ 0)) 
```

```{r fram graphing}
#transect start/stops for mapping
fram_lines <- tibble( x = c(-10, -20, -30, 0, -10, 0, 0, 0), 
                      y = c(0, 0, 0, 10, 50, 70, 80, 90), 
                      xend = c(-10, -20, -30, -33, -33, -60, -65, -39), 
                      yend = c(100, 100, 100, 10, 50, 70, 80, 90),
                      label = c("1NS", "2NS", "3NS", "1EW", "5EW", "7EW", "8EW", "9EW")
                    )


 fram_plot<-  ggplot( data = fram_1, mapping = aes(x = x, y = y)) +
  geom_point(size = 0.3) +
  geom_segment(data = fram_lines, aes(x = x, y = y, xend = xend, yend = yend)) +
   geom_text(data = fram_lines,
            aes( label = label),
            nudge_x = 4,
            size = 2,
            )

  # should expect to see mirroring of each point, originals & mirrors are colored 
  fram_plot +
    geom_point(data = filter(fram_1, zoned ==1), aes( colour = transect_orientation))+
    coord_fixed(ratio = 1, xlim = c(-75, 10), ylim = c(0, 100))
   # geom_line(aes( group = global_id), color = "red")

  fram_plot
  
  
fram_1 %>%
  dplyr::select(x,y,swapped_dir,object_id, transect_orientation)


  
  fram_1 %>%
    # filter(swapped_dir == 0) %>%
    mutate(swapped_dir = as.factor(swapped_dir)) %>%
ggplot(  aes(x,y) ) +
  geom_point(data = fram_1,  aes(x,y, shape = swapped_dir, 
                                 color = transect_orientation),
             size = 4) +
  geom_segment(data = fram_lines, aes(x = x, y = y, xend = xend, yend = yend)) +
    scale_shape_manual(values = c(16,1)) +
    xlim(c(-16.9,-23.1)) +
    ylim(c(46.9,53.1)) +
  geom_text(data = fram_1, aes(label = number_reproductive_in_group),
            size = 2) +
    coord_fixed(ratio = 1) +
    scale_colour_manual(values = c("yellow","skyblue")) +
  geom_text(data = fram_1, aes(label = object_id),
            nudge_x = 0.25,
            nudge_y = 0.25,
            size = 2.5) +
    coord_fixed(ratio = 1) 
  
fram_1 %>%
  filter(x > -23.1 & x < -16.9 &
           y > 46.9 & y < 53.1) %>%
  bind_rows(fram_1) %>%
  dplyr::select(x, y, object_id, transect_orientation, swapped_dir, number_reproductive_in_group, number_vegetative_in_group) %>%
  arrange(x, y)
  
```


```{r}

# only zoned, still some repeated object_ids
pic1Dups <- dup.fun(picnic_1[picnic_1$zoned == 1,], Tdistance = 1)

#Graphing the site with duplicated observations marked in blue

pic_lines <- tibble( x = c(-20, -30, -40, -20, -10, -10, 0), 
                     y = c(20, 10, 0, 20, 30, 40, 50), 
                     xend = c(-20, -30, -40, -90, -70, -70, -50),
                     yend = c(60, 76, 50, 20, 30, 40, 50),
                     label = c("1NS", "2NS", "3NS", "1EW", "2EW", "3EW", "4EW"))

duplicates <- pic1Dups %>%
  group_by(dupID) %>%
  filter(n() > 1)

# Select only the swapped when that location was closer
pic_plot <- pic1Dups %>% 
  filter(!(object_id %in% c(31,110,193,109, 115, 200) & is.na(swapped))) %>%
  # filter(zoned ==1 & swapped_dir == 0) %>% 
  ggplot( mapping = aes(x, y)) +
  geom_point(data = filter(picnic_1, swapped_dir == 0), aes(x,y), shape = 16, size = 1) +
  geom_point(aes(color = transect_orientation), shape = 1) +
  geom_point(data = duplicates, color = "white", size = 0.8) +
  scale_colour_manual("Orientation", values = c("#1E88E5","#D81B60"))  +
  geom_segment(data = pic_lines, aes(x = x, y = y, xend = xend, yend = yend)) +
  theme(
      axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),  #legend.position="none",
      panel.background = element_rect(fill='transparent'), #transparent panel bg
      plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
      panel.grid.major = element_blank(), #remove major gridlines
      panel.grid.minor = element_blank(), #remove minor gridlines
      legend.background = element_rect(fill='transparent'), #transparent legend bg
      legend.box.background = element_rect(fill='transparent') #transparent legend panel
    ) + 
  coord_fixed()

# FastUtils::getPlotDims(pic_plot)
# FastUtils::getPlotDims(pDensity2)
## See "DesignCoverage_DistanceSampling_sim.Rmd"
ggsave("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Sclerocactus-glaucus/Sclerocactus-glaucus_Manuscripts/DistanceSampling_Images/GraphicalAbstractBase.jpg",
pDensity2 +
  annotation_custom(grob = as.grob(pic_plot), xmin = 0.15, xmax = 0.87, ymin = 0.25, ymax = 0.87)
  ,
width=150, height=90,units='mm', dpi=300)


# pic_plot +
#   annotation_custom(grob = as.grob(pDensity2))
```


ids are duplicated in picnic_1 for the original and the swapped  
## Need to remove all the swapped ones that are not closer. 
group_by plant Id and only keep the one with the closer distance (if within the threshold) or only keep original.  
    case_when pair with other transect is closer and < 1 meter ~ keep this one
    case_when pair has neither swapped or original within 1 meter, keep original
    case_when direction_to_group was NA, keep the one with a closer distance to another one?? | if none within 1 meter, keep one at random. 


Mark-Recapture   
fixedDistDupID
```{r}
library(RMark)

## Should be 91 records out
picnic_1 %>%
  filter(zoned == 1 & swapped_dir == 0)

pic1Dups %>%
  group_by(object_id) %>%
  summarise(n = n()) 

pic1Dups %>%
  filter(object_id %in% c(31, 109, 110, 115, 193, 200)) %>%
  arrange(object_id)

x <- pic1Dups %>%
  # filter(!is.na(objectID)) %>% ## until I can fix this??
  mutate(dh = 1) %>%
  complete(transect_orientation, dupID, 
           fill = list(dh=0)) %>%
  pivot_wider(id_cols = dupID, values_from = dh, names_from = transect_orientation) %>%  
  unite(ch, EW:NS, sep="") 

## Only 84 total? 
x %>%
  filter(ch == "11")  ## 13 observed by both
x %>%
  filter(ch != "11")  ## 71 by only one
x %>%
  filter(nchar(ch) != 2) ## none

setwd("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScleroDistanceSampling")
x %>%
  mutate(ch = paste0(ch,";")) %>%
  write.table(file = "Data/Closed.inp",
              sep = " ", col.names = FALSE, row.names = FALSE)

scglClosed.inp <- convert.inp("Data/Closed.inp",
                              covariates = NULL,
                              use.comments = FALSE)

M <- nrow(x) ## (81) now 84 unique observations
clusterSize <- 1.58

## EW then NS
Cproc <- process.data(x, model = "Closed")
Cddl <- make.design.data(Cproc)

# constant p and c
closed <- mark(data = Cproc, Cddl, model.parameters = list(p = list(formula = ~1),
                                                           c = list(formula = ~1)))

closedshared <- mark(data = Cproc, Cddl, model.parameters = 
                       list(p = list(formula=~1,share=TRUE)))

get.real(closed, "p", se = TRUE)
get.real(closed, "f0", se = TRUE)
get.real(closedshared, "p", se = TRUE)
get.real(closedshared, "f0", se = TRUE) ## 93.72
out <- get.real(closedshared, "f0", se = TRUE)
out$estimate

## old: 121.98 (56.5 - 263.3) as the 95% CI of individuals never seen
(out$estimate+M)*clusterSize
(out$lcl + M)*clusterSize
(out$ucl + M)*clusterSize

```
I have lost 7 plants but otherwise very close!!


## MRDS  
```{r}

library(mrds)

picnic.data1 <- pic1Dups %>%
  filter(!(object_id %in% c(31,110,193,109, 115, 200) & is.na(swapped))) %>%
  ## rename distance_to_group to distance
  rename(distance = distance_to_group) %>%
  mutate(distance = abs(distance)) %>%
  mutate(detected = 1) %>%
  rename("object" = dupID) %>%
  rename("observer" = transect_orientation) %>%
  rowwise() %>%
  mutate(size = sum(c(number_seedlings_in_group, 
                      number_vegetative_in_group, 
                      number_reproductive_in_group), na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(object) %>%
  mutate(size = max(size, na.rm = TRUE)) %>%
  ungroup() %>%
  complete(nesting(object,size),observer, fill = list(detected = 0)) %>% ## Keep only size and plant ID that are in the data
  dplyr::select(object,observer, size, detected, distance, transect_meter_mark) %>%
  arrange(object, observer) %>%
  mutate(observer = as.factor(as.numeric(as.factor(observer)))) %>%
  relocate(size, .after = distance) %>%
  mutate(object = as.integer(as.factor(object))) %>%
  ## When grouped, one might be missed, that's the row where distance missing
  group_by(object) %>%
  ## each set of two observations
  ## need to update the group size as well. 
  mutate(distance2 = case_when(
    detected == 0 & max(abs(transect_meter_mark), na.rm = TRUE) < 15 ~ abs(10-max(abs(transect_meter_mark), na.rm = TRUE)),
    detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 15 & max(abs(transect_meter_mark), na.rm = TRUE) < 25 ~ 
                                                                             abs(20-max(abs(transect_meter_mark), na.rm = TRUE)),
    detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 25 & max(abs(transect_meter_mark), na.rm = TRUE) < 35 ~ 
                                                                             abs(30-max(abs(transect_meter_mark), na.rm = TRUE)),
    detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 35 & max(abs(transect_meter_mark), na.rm = TRUE) < 45 ~ 
                                                                             abs(40-max(abs(transect_meter_mark), na.rm = TRUE)),
    detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 45 & max(abs(transect_meter_mark), na.rm = TRUE) < 55 ~ 
                                                                             abs(50-max(abs(transect_meter_mark), na.rm = TRUE)),
    detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 55 & max(abs(transect_meter_mark), na.rm = TRUE) < 65 ~ 
                                                                             abs(60-max(abs(transect_meter_mark), na.rm = TRUE)),
      TRUE ~ max(distance, na.rm = TRUE)
                              )) %>%
  ungroup() %>%
  dplyr::select(object:detected,size,distance2) %>%
  rename("distance" = "distance2") %>%
  group_by(object,observer) %>%
  ## Where the size differs, only keep larger estimate of size; won't see more than are there, might miss some
  filter(size == max(size, na.rm = TRUE)) %>%
  ungroup() %>%
  distinct()
  
picnic.data1 %>%
  group_by(observer, detected) %>%
  summarise(n = n())

picnic.data1 %>%
  group_by(object) %>%
  filter(detected == 1) %>%
  summarise(n = n()) %>%
  filter(n > 1)

sizeVariation <- table(picnic.data1$object) ## A few have duplicate observations! so four instead of 2. Fixed by selecting larger
picnic.data1 %>%
  filter(object %in% c(sizeVariation != 2))

## Fully independent   
# MR only, no distance data
fi.mr.dist <- ddf(method = "trial.fi", 
                  mrmodel = ~ glm(link = 'logit', formula = ~ size), #  distance),
                  data = picnic.data1,
                  meta.data = list(width = 3))


detection.tables <- det.tables(fi.mr.dist)
# Print these detection tables
print(detection.tables)
plot(fi.mr.dist)

picnic.data1 %>%
  ggplot(  aes(distance, fill = observer)) +
    geom_histogram()


fi.mr.dist.io <- ddf(method = "io.fi", 
                  mrmodel = ~ glm(link = 'logit', formula = ~ size), #  distance),
                  data = picnic.data1,
                  meta.data = list(width = 3))


detection.tables.io <- det.tables(fi.mr.dist.io)
# Print these detection tables
print(detection.tables.io)
plot(fi.mr.dist.io)

############ Swap observer 1 and 2
picnic.data1.swapObs <- picnic.data1 %>%
  mutate(observer = case_when(observer == 1 ~ 2,
                              observer == 2 ~ 1))

fi.mr.dist.swap <- ddf(method = "trial.fi", 
                  mrmodel = ~ glm(link = 'logit', formula = ~ size), #  distance),
                  data = picnic.data1.swapObs,
                  meta.data = list(width = 3))


detection.tables.swap <- det.tables(fi.mr.dist.swap)
# Print these detection tables
print(detection.tables.swap)
plot(fi.mr.dist.swap)

```

From AnalysisCrossedDS.Rmd
```{r}

## Scale the abundance by the detection rate
# 2244 (1877 - 2632)
2244/0.5 ## trial.fi
2244/0.46 ## io.fi
2244/0.27 ## closed mark-recapture


```

Distance for CV 

```{r}
library(Distance)

picnic.region <- data.frame(Region.Label = c("EW","NS"), 
                            Area = 7802)
picnic.units <- data.frame(Variable = c("Area","Effort","distance"),
                           Units = c("square kilometre", "meter","meter"),
                           conversion = c(0.000001, 1, 1))

picnic.samples <- pic1Dups %>%
  filter(!(object_id %in% c(31,110,193,109, 115,200) & is.na(swapped))) %>%
  distinct(transect_number, transect_orientation, transect_length, object_id, distance_to_group) %>%
  rowwise() %>%
  mutate(Sample.Label = paste0(transect_number,transect_orientation)) %>%
  ungroup() %>%
  mutate(Region.Label = transect_orientation) %>%
  rename("Effort" = "transect_length") %>%
  # dplyr::select(Sample.Label, Region.Label, Effort) %>%
  mutate(Effort = case_when(Sample.Label == "1EW" ~ 70,
                            TRUE ~ Effort)) %>%
  filter(!is.na(Effort)) %>%
  arrange(Sample.Label) %>%
  mutate(Area = 7802) %>%
  rename(distance = distance_to_group) %>%
  rename(object = object_id) %>%
  dplyr::select(Region.Label, Area, Sample.Label, Effort, object, distance) %>%
  mutate(Study.Area = "Picnic") %>%
  mutate(distance = abs(distance))

picnic.hn <- ds(data = picnic.samples, key = "hn", convert_units = convert_units("meter","meter","square meter"))
summary(picnic.hn)
print(picnic.hn$dht$individuals$D)
### cv is EW: 0.166; NS: 0.168

cvs <- data.frame(Strata = c("EW","NS","total"), 
                    meanD = picnic.hn$dht$individuals$D$Estimate,
                    cvD = picnic.hn$dht$individuals$D$cv)

                    



pic1Dups %>%
  filter(!(object_id %in% c(31,110,193,109, 115, 200) & is.na(swapped))) %>%
  group_by(object_id) %>%
  summarise(n = n()) %>%
  filter(n != 1)


```

