---
title: "2026_SCGL_work"
author: "Jonathan Amegin"
date: "2026-05-01"
output: html_document
editor_options: 
  markdown: 
    wrap: 90
---

Updates by Michelle DePrenger-Levin 2026.06.02

To do:

~~- for each observation, create clones in the data frame that switch the
"direction_from_transect" field for later duplicate analysis~~

~~- Split the comparison of group sizes up between seedlings, vegetative, and
reproductive~~

~~-   Develop probability scoring to show 'percent similarity' between observations based on
    weighted scores of covariates

~~-   Keep only the duplicates with the shortest distance between the compared observations
    so that only 1 duplicate/global_id

##Setting Up

```{r message=FALSE, warning=FALSE, include=FALSE}
# tidyverse contains both 'dyplr' and 'ggplot2'
library(tidyverse)
library(janitor)
library(uuid)

setwd("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Sclerocactus-glaucus/Sclerocactus-glaucus_RawData/2024_CrossedDistanceSampling")
#this contains all observations of clusters of SCGL for all sites & Transects
scgl_obs <- read_csv("distanceSampling_1.csv")
scgl_obs <- clean_names(scgl_obs)

#This contains metadata for each transect including length & direction
scgl_trans_length <- read_csv("transect_2.csv")
scgl_trans_length <- clean_names(scgl_trans_length)

#filtering NA's and outliers, and changing some variable classes for later work
scgl <- scgl_obs[!(is.na(scgl_obs$transect_meter_mark|scgl_obs$distance_to_group)),] %>%  
        filter(transect_meter_mark < 110, distance_to_group < 10) %>% 
        mutate( object_id = as.factor(object_id),
                site_name = as.factor(site_name),
                transect_number = as.factor(transect_number),
                swapped_dir = 0)

scgl[is.na(scgl[,"number_seedlings_in_group"]), "number_seedlings_in_group"] <- 0
scgl[is.na(scgl[,"number_vegetative_in_group"]), "number_vegetative_in_group"] <- 0
scgl[is.na(scgl[,"number_reproductive_in_group"]), "number_reproductive_in_group"] <- 0

#adding clone observations with the 'direction_to_group' swapped for later analysis
#some observations had NA, so need to check both directons for them
### The NA+ and NA-  see lines 74, default is + for direction
scgl1 <- rbind(scgl,
               mutate(scgl,
                      swapped_dir = "1",
                      direction_to_group = case_when(direction_to_group == "+" ~ "-",
                                                     direction_to_group == "-" ~ "+",
                                                     is.na(direction_to_group) ~ "NA-"
                                                     )))

scgl1[is.na(scgl1[,"direction_to_group"]), "direction_to_group"] <- "NA+"

#use the attributes to add modifiers to the measurements
#EW transect meter marks need to be * -1 because they were recorded as if in the(+,+) quadrant, but the mapping we will do is based on (-,+) quadrant.             
scgl1 <-
  scgl1 %>%
  mutate(
    transect_meter_mark = case_when(
      transect_orientation == "EW" ~ transect_meter_mark * -1,
      TRUE ~ transect_meter_mark),
    #making the 'direction to group' indicate if the 'distance to group' needs to be left/right of transect.  
    distance_to_group = case_when(
      direction_to_group == "-" | direction_to_group == "NA-" ~ distance_to_group * -1,
                                                         TRUE ~ distance_to_group
      ))
           
#setting up the (X,Y) columns to then transform based on plot/shapefile analysis
scgl1 <- scgl1 %>% 
  mutate( x = case_when(transect_orientation == "NS" ~ distance_to_group,
                        transect_orientation == "EW" ~ transect_meter_mark), 
          y = case_when(transect_orientation == "EW" ~ distance_to_group,
                        transect_orientation == "NS" ~ transect_meter_mark)) 

#joining the two dataframes to add transect metadata to each observation
scgl1 <- left_join(scgl1, scgl_trans_length[, c("transect_length", "global_id")],
                   by = join_by(parent_global_id == global_id))                
                
#filtering to sites
fram <- scgl1 %>% 
  filter(site_name == "12")

picnic <- scgl1 %>% 
  filter(site_name == "1"| site_name == "3")

bridgeport <- scgl1 %>% 
  filter(site_name == "10")

oilpad <- scgl1 %>% 
  filter(site_name == "11")

```


### 2026 function

Idea is to use covariates to generate a probablity score indicating likelihood of pairs of
observations being duplicates.

Covariates affecting probability:

-   Number in groups, split between categories.

-   Distance between observations

-   Direction of transect

-   within/out the crossed area

function to apply comparison to each object_id pair by referencing the fram_1 table Needs
to use the combn() pairs and parse the fram_1 table and then compare the covariates to
evaluate a probability score

```{r}
###Custom Function 1
dup.fun <- function(site, distance, cluster, TS = 8, TV = 5, TR = 2){
  
  #making new column with new UUID for each observation due to cloned entries
  site$new_id = UUIDgenerate(use.time = FALSE, n = nrow(site), output = "string")
  
  #Creating list of all combinations of global_id's to reference
  list <- utils::combn(site$new_id, m = 2, simplify = FALSE)
  #Appending calculated variables to the list  
  list <- lapply(list, function(list.1){

    #indexing data with the paired global_ids
    a <- site[site$new_id == list.1[1],]
    b <- site[site$new_id == list.1[2],]

    #Calculating distance
    list.1[3] <- dist(rbind(c(a$x,a$y), c(b$x,b$y)), method = "euclidean")

    #determining if they are on the same transect
    ## Then take out the ones that are on the same
    list.1[4] <- case_when(a$transect_orientation == b$transect_orientation ~ 1,
                           a$transect_orientation != b$transect_orientation ~ 0)

    #Group size comparisons
    list.1[5] <- abs(diff(c(a$number_seedlings_in_group,b$number_seedlings_in_group)))
    list.1[6] <- abs(diff(c(a$number_vegetative_in_group,b$number_vegetative_in_group)))
    list.1[7] <- abs(diff(c(a$number_reproductive_in_group,b$number_reproductive_in_group)))
    
    #Comparing if the 'direction_from_transect' variable was swapped earlier
    #attempts to correct for potential recorder error
    #could change to a calculation to make weighting the probability score easier
    ## Michelle notes: (1) check if the swapped point is closer than the original
    list.1[8] <- as.numeric(a$swapped_dir) + as.numeric(b$swapped_dir) ## sums 0, 1, or 2 to see who was swapped
    list.1[9] <- case_when(str_detect(a$direction_to_group, pattern = "NA") &
                            str_detect(b$direction_to_group, pattern = "NA") ~ 2,
                           xor(str_detect(a$direction_to_group, pattern = "NA"),  ## if one is true == xor
                            str_detect(b$direction_to_group, pattern = "NA")) ~ 1,
                            TRUE ~ 0)
    #The orientation of id_1
    list.1[10] <- a$transect_orientation
    list.1[11] <- b$transect_orientation
     return(list.1)
  })

  
  #use the Hamilton, et. al discrepancy score equation as a base
  ## only want ones on different transects (keep if want to follow Hamilton and compare known non-duplicates)
  ## Larger discrepancy score == less likely to be a duplicate
  ## Numbers don't matter (picked arbitrarily in their appendix 2a) 
  list <- lapply(list, function(z){
    ## Discrepancy score
    z[12] <- 
      # if distance between is greater than the set distance > 1, if less than, is under 1
      # (min(as.numeric(z[3]), distance)/ distance) + # using distance in meters as the 'beyond a doubt' not duplicate distance
      # (min(as.numeric(z[5]), TS)/TS) + #seedlings difference, 8 = not duplicate; TS
      # (min(as.numeric(z[6]), TV)/TV) + #vegetative difference, 5 = not duplicate; TV
      # (min(as.numeric(z[7]), TR)/TR) #reproductive difference, 2 = not duplicate; TR # to give margin 
      ## Remove min because will either be 1 or less than one
      (as.numeric(z[3])/ distance) +
      (as.numeric(z[5])/TS) + #seedlings difference, 8 = not duplicate; TS
      (as.numeric(z[6])/TV) + #vegetative difference, 5 = not duplicate; TV
      (as.numeric(z[7])/TR) #reproductive difference, 2 = not duplicate; TR # to give margin 
    
  ## Fixed tolerance, just by distance and only total group size, cluster is variable you can put in. 
    z[13] <-  #fixed tolerance test
      # if((as.numeric(z[3]) < distance) & (sum(as.numeric(z[5],z[6],z[7])) < cluster)) 1 else 0
      if(( as.numeric(z[3]) < distance ) & 
         ( as.numeric(z[5]) < TS ) & ( as.numeric(z[6]) < TV ) & ( as.numeric(z[6]) < TR )) 1 else 0
    
    
    z[14] <-  #fixed tolerance test based only on distance
      if((as.numeric(z[3]) < distance) ) 1 else 0

    return(z)
  })
  
    df <- data.frame(do.call(rbind,list))
    names(df) <- c("id_1", "id_2",
                      "distance",
                      "same_diff",  ## if on same or different transect 1 = same, 0 = different
                      "diff_seedlings",
                      "diff_vegetative",
                      "diff_reproductive",
                      "swapped_dir",
                      "NA_dir",
                      "orientation_1",
                      "orientation_2",
                      "discrepency",
                      "fixed_tolerance",
                      "fixed_disttolerance")
    df <- df %>%
      mutate("ThresholdSeedling" = TS, "ThresholdVeg" = TV,"ThresholdRep" = TR) %>%
      mutate_at(vars(distance:NA_dir, discrepency:ThresholdRep), as.numeric)
  
 df$dup_id <- UUIDgenerate(use.time = FALSE, n = nrow(df), output = "string") # A new ID to join and mutate to add this same ID to both; Jonathan will add this! 
 
 ## df wide to long splitting transect orientation and id 
 df_long <- df %>%
   pivot_longer(starts_with("id"),
                values_to = "id") %>%
   dplyr::select(distance:dup_id, id) %>%
   pivot_longer(starts_with("orientation"),
                values_to = "trans_orientation", 
                names_prefix = "orientation_") %>%
   dplyr::select(distance:id, trans_orientation) %>%
   distinct()

 ## Need to have assigned duplicate IDs for (1) fixed distance, (2) fixed distance and cluster size/stage, (3) probabilistic inverse logit
 # Pull all new_ids that would be duplicates by fixed distance
 dup_fixedDist <- df_long[df_long$same_diff == 0 & df_long$fixed_disttolerance == 1, c("id","dup_id")] %>%
   distinct()
 dup_fixedDistClust <- df_long[df_long$same_diff == 0 & df_long$fixed_tolerance == 1, c("id","dup_id","diff_reproductive","distance")] %>%
   distinct() %>%
   group_by(id) %>%
   filter(distance == min(distance))
 discrepScore <- df_long[df_long$same_diff == 0 & df_long$discrepency < 1, c("id","dup_id","discrepency")] %>%
   rowwise() %>%
   mutate(discrepYN =  rbinom(1,1,prob = discrepency)) %>%
   filter(discrepYN == 0) %>% 
   distinct() %>%
   group_by(id) %>%
   filter(discrepency == min(discrepency))

 testSite <- site %>%
   full_join(dup_fixedDist, by = c("new_id" = "id")) %>%
   mutate(fixedDistDupID = case_when(is.na(dup_id) ~ new_id,
                                     TRUE ~ dup_id)) %>%
   dplyr::select(fixedDistDupID, distance_to_group, new_id, transect_orientation, swapped_dir,
                 number_seedlings_in_group, number_vegetative_in_group, number_reproductive_in_group,
                 transect_meter_mark) %>%
   mutate(distance_to_group = abs(distance_to_group)) %>% ## Distance is already the distance between points
   full_join(discrepScore, by = c("new_id" = "id")) %>%
   mutate(discrepDupId = case_when(is.na(dup_id) ~ new_id,
                                     TRUE ~ dup_id)) %>%
   ## dup_id isn't matching .x and .y because it either selects the dup_id or keeps the new_id
   dplyr::select(fixedDistDupID, discrepDupId, new_id, distance_to_group:transect_meter_mark, 
                 discrepency:discrepYN) %>%
   full_join(dup_fixedDistClust, by = c("new_id" = "id")) %>%
   mutate(fixedDistClustDupID = case_when(is.na(dup_id) ~ new_id,
                                     TRUE ~ dup_id)) %>%
   dplyr::select(fixedDistDupID,fixedDistClustDupID,discrepDupId,new_id, dup_id, distance_to_group,
                 transect_orientation:discrepYN, diff_reproductive, distance) 
   
 
 return(list(testSite, df, df_long))
}

```



```{r}
####Custom Function 2
#checking for observations that have multiple duplicates & then filtering down to 'most likely'
## If discrepancy is over 1, then for sure isn't a duplicate! Filters down to the ones that 'could' be 
## fixed_tolerance == 1: cluster size and distance is under tolerance, 0: over tolerance
## add a fixed_tolerance for just distance
dup.fun.1 <-
  function(site){
  df <- filter(site, discrepency <1 & orientation == 0) #filter to possible dupes, removed known not duplicates
  id <- c(df$id_1, df$id_2) #vector of all id's that are associated with duplicated observations
  if (sum(duplicated(id)) != 0){
    dup.id <- unique(id[duplicated(id)])
    id.1 <- #function to find the multiples that are NOT the most likely
    lapply(dup.id, function(i){
        df.1 <- filter(df, id_1 == i | id_2 == i)
        df.2 <- filter(df.1, discrepency != min(discrepency))$dup_id
        df.2
    })
    df[-which(df$dup_id %in% unlist(id.1)),]}
  else{df} 
}

###Custom Function 3
# Getting the numbers needed for the unmakred package: Unmakred dulicate observation function
# This function calculates the number of observations made by each observer (different transect directions) 
# alongside the number of observations said to be duplicates based on the earlier dup.fun function
# and outputs a named list
## Filters to the minimum discrepancy score

# Use only original or also swapped: swap is true and false (use swapped or not)
# 
umdo.fun <- 
function(site, swap, d, c){
  #allows choice of using the cloned/mirrored observations for checking errors
  if(swap == FALSE){ 
  site.1 <- filter(site, swapped_dir == 0 & zoned == 1) # only the zoned ones
  site.ew <- filter(site.1, transect_orientation == "EW")
  site.ns <- filter(site.1, transect_orientation == "NS")
  site.1d <- dup.fun(site.1, distance = d, cluster = c)
  #getting the numbers of duplicates for both the discrepancy score & fixed tolerance 
  #analysis allows for comparison between methods
  site.score <- dup.fun.1(filter(site.1d, discrepency < 1))
  site.fixed <- dup.fun.1(filter(site.1d, fixed_tolerance == 1))
  # site.fdistance <- 
  
  
  list <- list("EW" = nrow(site.ew), 
               "NS" = nrow(site.ns),
               "scored" = nrow(site.score),
               "fixed" = nrow(site.fixed))
  }else{
  site.1 <- filter(site, zoned == 1)
  site.ew <- filter(site.1, transect_orientation == "EW" & swapped_dir == 0)
  site.ns <- filter(site.1, transect_orientation == "NS" & swapped_dir == 0)
  site.1d <- dup.fun(site.1, distance = d, cluster = c)
  #getting the numbers of duplicates for both the discrepency score & fixed tolerance 
  #analysis allows for comparison between methods
  site.score <-   dup.fun.1(filter(site.1d, discrepency < 1 & swapped_dir == 0)) # neither swapped
  site.score.1 <- dup.fun.1(filter(site.1d, discrepency < 1 & swapped_dir < 2)) # one swapped
  site.score.2 <- dup.fun.1(filter(site.1d, discrepency < 1 & swapped_dir <= 2)) # two or one swapped
  site.fixed <-   dup.fun.1(filter(site.1d, fixed_tolerance == 1 & swapped_dir == 0))
  site.fixed.1 <- dup.fun.1(filter(site.1d, fixed_tolerance == 1 & swapped_dir < 2))
  site.fixed.2 <- dup.fun.1(filter(site.1d, fixed_tolerance == 1 & swapped_dir <= 2))
  ## and fixed distance
  
  
  list <- list("EW" = nrow(site.ew), 
               "NS" = nrow(site.ns),
               "scored" = nrow(site.score),
               "scored swap 1" = nrow(site.score.1),
               "scored swap 2" = nrow(site.score.2),
               "fixed" = nrow(site.fixed),
               "fixed swap 1" = nrow(site.fixed.1),
               "fixed swap 2" = nrow(site.fixed.2))
  }
  return(list)
}


```



##Fram

###Plotting individuals

Transforming Fram coordinates to accurately reflect plot design.This is based on the
origin (0,0) being in the lower right of the plot so that one or both of the axis matches
the start of at least one transect. In this case, the origins of the three NS transects
are at Y = 0, and 4 out of 5 EW transects' origins are at X = 0.

```{r fram transform}

fram_1 <- fram %>% 
  mutate( 
    x = case_when(transect_orientation == "NS" & transect_number == "1" ~ x - 10, 
                  transect_orientation == "NS" & transect_number == "2" ~ x - 20,
                  transect_orientation == "NS" & transect_number == "3" ~ x - 30,
                  transect_orientation == "EW" & transect_number == "5" ~ x - 10,
                  TRUE ~ x
                  ),
    y = case_when(transect_orientation == "EW" & transect_number == "1" ~ y + 10,
                  transect_orientation == "EW" & transect_number == "5" ~ y + 50,
                  transect_orientation == "EW" & transect_number == "7" ~ y + 70,
                  transect_orientation == "EW" & transect_number == "8" ~ y + 80,
                  transect_orientation == "EW" & transect_number == "9" ~ y + 90,
                  TRUE ~ y),
    )

#Fram NS lines tapes were laid backwards making 0 = 100. 
fram_1 <- fram_1 %>% 
  mutate( y = case_when(transect_orientation =="NS" ~ ((y - 100)* -1), TRUE ~ y)
    )

fram_1 <- mutate(fram_1, zoned = 
                   case_when(
                     c(rowSums(mapply( FUN = function(a, b){between(a, b-3, b+3)}, fram_1[, "y"], c(10, 50, 70, 80, 90))) +
                                                rowSums(mapply( FUN = function(a, b){between(a, b-3, b+3)}, fram_1[, "x"], c(-10, -20, -30)))) == 2 ~ 1,
       TRUE ~ 0)) 
```

```{r fram graphing}
#transect start/stops for mapping
fram_lines <- tibble( x = c(-10, -20, -30, 0, -10, 0, 0, 0), 
                      y = c(0, 0, 0, 10, 50, 70, 80, 90), 
                      xend = c(-10, -20, -30, -33, -33, -60, -65, -39), 
                      yend = c(100, 100, 100, 10, 50, 70, 80, 90),
                      label = c("1NS", "2NS", "3NS", "1EW", "5EW", "7EW", "8EW", "9EW")
                    )


 fram_plot<-  ggplot( data = fram_1, mapping = aes(x = x, y = y)) +
  geom_point(size = 0.3) +
  geom_segment(data = fram_lines, aes(x = x, y = y, xend = xend, yend = yend)) +
   geom_text(data = fram_lines,
            aes( label = label),
            nudge_x = 4,
            size = 2,
            )

  # should expect to see mirroring of each point, originals & mirrors are colored 
  fram_plot +
    geom_point(data = filter(fram_1, zoned ==1), aes( colour = transect_orientation))+
    coord_fixed(ratio = 1, xlim = c(-75, 10), ylim = c(0, 100))
   # geom_line(aes( group = global_id), color = "red")

  fram_plot
  
  
fram_1 %>%
  dplyr::select(x,y,swapped_dir,object_id, transect_orientation)


  
  fram_1 %>%
    # filter(swapped_dir == 0) %>%
    mutate(swapped_dir = as.factor(swapped_dir)) %>%
ggplot(  aes(x,y) ) +
  geom_point(data = fram_1,  aes(x,y, shape = swapped_dir, 
                                 color = transect_orientation),
             size = 4) +
  geom_segment(data = fram_lines, aes(x = x, y = y, xend = xend, yend = yend)) +
   # geom_text(data = fram_lines, 
   #          aes( label = label),
   #          nudge_x = 4,
   #          size = 2,
   #          ) +
    scale_shape_manual(values = c(16,1)) +
    xlim(c(-16.9,-23.1)) +
    ylim(c(46.9,53.1)) +
  geom_text(data = fram_1, aes(label = number_reproductive_in_group),
            size = 2) +
    coord_fixed(ratio = 1) +
    scale_colour_manual(values = c("yellow","skyblue")) +
  geom_text(data = fram_1, aes(label = object_id),
            nudge_x = 0.25,
            nudge_y = 0.25,
            size = 2.5) +
    coord_fixed(ratio = 1) 
  
fram_1 %>%
  filter(x > -23.1 & x < -16.9 &
           y > 46.9 & y < 53.1) %>%
  bind_rows(fram_1) %>%
  mutate()
  dplyr::select(x, y, object_id, transect_orientation, swapped_dir, number_reproductive_in_group, number_vegetative_in_group) %>%
  arrange(x, y)
  
```

Testing
```{r}

site <- fram_1
#making new column with new UUID for each observation due to cloned entries
  site$new_id = UUIDgenerate(use.time = FALSE, n = nrow(site), output = "string")
  #Creating list of all combinations of global_id's to reference
  list <- utils::combn(site$new_id, m = 2, simplify = FALSE)
  #Appending calculated variables to the list  
  list <- lapply(list, function(list.1){

    #indexing data with the paired global_id's
    a <- site[site$new_id == list.1[1],]
    b <- site[site$new_id == list.1[2],]

    #Calculating distance
    list.1[3] <- dist(rbind(c(a$x,a$y), c(b$x,b$y)), method = "euclidean")

    #determining if they are on the same transect
    list.1[4] <- case_when(a$transect_orientation == b$transect_orientation ~ 1,
                           a$transect_orientation != b$transect_orientation ~ 0)

    #Group size comparisons
    list.1[5] <- abs(diff(c(a$number_seedlings_in_group,b$number_seedlings_in_group)))
    list.1[6] <- abs(diff(c(a$number_vegetative_in_group,b$number_vegetative_in_group)))
    list.1[7] <- abs(diff(c(a$number_reproductive_in_group,b$number_reproductive_in_group)))
    
    #Comparing if the 'direction_from_transect' variable was swapped earlier
    #attempts to correct for potential recorder error
    #could change to a calculation to make weighting the probability score easier
    list.1[8] <- as.numeric(a$swapped_dir) + as.numeric(b$swapped_dir)
    list.1[9] <- case_when(str_detect(a$direction_to_group, pattern = "NA") &
                            str_detect(b$direction_to_group, pattern = "NA") ~ 2,
                           xor(str_detect(a$direction_to_group, pattern = "NA"), 
                            str_detect(b$direction_to_group, pattern = "NA")) ~ 1,
                            TRUE ~ 0)
     return(list.1)
  })
  
all.fram <- do.call(rbind, list) %>%
  as.data.frame() %>%
  filter(V3 < 1)
```



Looks like EW 291 matches NS 340 when EW 291 IS NOT swapped but NS 340 IS swapped     
Looks like EW 290 matches NS 


```{r}
#This is just testing some things out, safe to ignore 
test.fram <- dup.fun(filter(fram_1, zoned == 1), distance = 0.25, cluster = 5)

filter(test.fram, swapped_dir == 0)$discrepency %>% 
  hist(labels = TRUE)

nrow(filter(test.fram, discrepency < 0.5))

sapply(seq(0,1,0.05), function(i){
  nrow(filter(test.fram,  swapped_dir == 0 & discrepency < i))
})

umdo.fun(fram_1, swap = FALSE)
umdo.fun(fram_1, swap = TRUE)


  # filter(test.fram, distance <8, discrepency <1)$distance %>% 
  # hist( breaks = 30)
```



##picnic
```{r}
#transforming coordinates to match the plot design. This requires accurate note taking
 picnic_1 <-  picnic %>% 
   mutate( 
    x = case_when(transect_orientation == "NS" & transect_number == "1" ~ x - 20, 
                  transect_orientation == "NS" & transect_number == "2" ~ x - 30,
                  transect_orientation == "NS" & transect_number == "3" ~ x - 40,
                  transect_orientation == "EW" & transect_number == "1" ~ x - 20,
                  transect_orientation == "EW" & transect_number == "2" ~ x - 10,
                  transect_orientation == "EW" & transect_number == "3" ~ x - 10,
                  transect_orientation == "EW" & transect_number == "4" ~ x
                  ),
    y = case_when(transect_orientation == "NS" & transect_number == "1" ~ y + 20, 
                  transect_orientation == "NS" & transect_number == "2" ~ y + 10,
                  transect_orientation == "NS" & transect_number == "3" ~ y ,
                  transect_orientation == "EW" & transect_number == "1" ~ y + 20,
                  transect_orientation == "EW" & transect_number == "2" ~ y + 30,
                  transect_orientation == "EW" & transect_number == "3" ~ y + 40,
                  transect_orientation == "EW" & transect_number == "4" ~ y + 50,
                  ))

#adding an attribute to show if an observation is within the zone that NS & EW transects share
picnic_1 <-
  mutate(picnic_1,
         zoned = case_when(
           c(rowSums(mapply(FUN = function(a, b){between(a, b-3, b+3)}, picnic_1[, "y"], c(20, 30, 40, 50))) +
             rowSums(mapply(FUN = function(a, b){between(a, b-3, b+3)}, picnic_1[, "x"], c(-20, -30, -40)))) == 2 ~ 1,
           TRUE ~ 0))


# umarked.picnic <- umdo.fun(picnic_1, swap = TRUE, d = 1, c = 5)

# [[1]]: the input site with the unique ID (new_id) added and the duplicates identified according to fixed distance, fixed distance and cluster size cutoffs, and rbinom for discrepency scores under 1 (over 1 considered not a duplicate)
# [[2]]: the data frame of paired comparisons
# [[3]]: the data frame in long form with unique values (instead of comparison of 1 vs 2 and 2 vs 1 in the wide data frame)
pic1Dups <- dup.fun(picnic_1[picnic_1$zoned == 1,], distance = 1, cluster = 4, TS = 8, TV = 3, TR = 2)
pic1Dups_821 <- dup.fun(picnic_1[picnic_1$zoned == 1,], distance = 0.5, cluster = 10, TS = 8, TV = 2, TR = 1)
## Higher discrepancy score == less likely to be a duplicate. Lower discrepancy score (approaching zero) == more likely duplicate
# #4 "orientation" 1 == same transect, 0 == different transect orientation
# #5 "diff_seedlings" diff seedlings; #6 diff veg; #7 diff repro
# #10 is a sum of how much distance is less than threshold +
#                 how much seedling + veg + rep is less than threshold

# Only slightly lower discrepancy score for those on different grid lines. Not surprising 
### 1 == same transect (known NOT duplicates)
### 0 == different transects (possible duplicates)
pic1Dups[[2]] %>% 
  filter(fixed_tolerance == 1) %>%
  ggplot(  aes(discrepency, fill = as.factor(same_diff)))+
    geom_density(alpha = 0.5) +
    theme_bw() 

pic1Dups[[2]] %>% 
  filter(fixed_disttolerance == 1) %>%
  ggplot(  aes(discrepency, fill = as.factor(same_diff)))+
    geom_density(alpha = 0.5) +
    theme_bw() 

# distance and cluster size
pic1Dups_821[[2]] %>% 
  filter(fixed_tolerance == 1) %>%
  ggplot(  aes(discrepency, fill = as.factor(same_diff)))+
    geom_density(alpha = 0.5) +
    theme_bw()

pic1Dups[[2]] %>%
  rowwise() %>%
  mutate(discrepYN = rbinom(1,1, prob = discrepency)) %>%
  filter(discrepYN == 0) %>%
  ggplot(  aes(discrepency, fill = as.factor(same_diff)))+
    geom_density(alpha = 0.5) +
    theme_bw()

pic1Dups[[2]] %>%
  filter(discrepency < 1)
## 
pic1Dups[[2]] %>%
  filter(discrepency < 1) %>%
  rowwise() %>%
  mutate(discrepYN = rbinom(1,1, prob = discrepency)) %>%
  # filter(same_diff == 0) %>%
  ggplot(  aes(distance, discrepYN, color = as.factor(same_diff))) +
    geom_point(pch = "|")+
    stat_smooth(method="glm",  method.args = list(family = "binomial"), se=TRUE) +
    theme_bw() +
    scale_color_discrete("Transect", labels = c("Possible duplicates","Known NON-duplicates"))

pic1Dups[[2]] %>%
  filter(discrepency < 1) %>%
  rowwise() %>%
  mutate(discrepYN = rbinom(1,1, prob = discrepency)) %>%
  # filter(same_diff == 0) %>%
  ggplot(  aes(distance, discrepYN, color = as.factor(same_diff))) +
    geom_point(pch = "|")+
    stat_smooth(method="glm",  method.args = list(family = "binomial"), se=TRUE) +
    theme_bw() +
    scale_color_discrete("Transect", labels = c("Possible duplicates","Known NON-duplicates"))

## Zero are the duplicates
pic1Dups[[2]] %>%
  filter(discrepency < 1) %>%
  rowwise() %>%
  mutate(discrepYN = rbinom(1,1, prob = discrepency)) %>%
  filter(same_diff == 0) %>%
  ggplot(  aes(distance, diff_reproductive)) +
    geom_point()+
    theme_bw() +
    facet_wrap(~ discrepYN)


```

ids are duplicated in picnic_1 for the original and the swapped  
## Need to remove all the swapped ones that are not closer. 
group_by plant Id and only keep the one with the closer distance (if within the threshold) or only keep original.  
    case_when pair with other transect is closer and < 1 meter ~ keep this one
    case_when pair has neither swapped or original within 1 meter, keep original
    case_when direction_to_group was NA, keep the one with a closer distance to another one?? | if none within 1 meter, keep one at random. 
```{r}
picnic_1 %>%
  ggplot(  aes(object_id)) +
    geom_histogram(stat = "count") +
    facet_wrap(~ swapped_dir)
```


Mark-Recapture   
fixedDistDupID
```{r}
# only zoned
pic1Dups <- dup.fun(picnic_1[picnic_1$zoned == 1,], distance = 1, cluster = 13, TS = 8, TV = 3, TR = 2)

x <- pic1Dups[[1]] %>%
  mutate(dh = 1) %>%
  complete(transect_orientation, fixedDistDupID, 
           fill = list(dh=0)) %>%
  pivot_wider(id_cols = fixedDistDupID, values_from = dh, names_from = transect_orientation) %>%  
  unite(ch, EW:NS, sep="") 

x %>%
  filter(ch == "11")

## 66 observed by both

setwd("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScleroDistanceSampling")
x %>%
  mutate(ch = paste0(ch,";")) %>%
  write.table(file = "Data/Closed.inp",
              sep = " ", col.names = FALSE, row.names = FALSE)

scglClosed.inp <- convert.inp("Data/Closed.inp",
                              covariates = NULL,
                              use.comments = FALSE)

M <- nrow(x) ## 81 unique observations
clusterSize <- 1.58

## EW then NS
Cproc <- process.data(x, model = "Closed")
Cddl <- make.design.data(Cproc)

# constant p and c
closed <- mark(data = Cproc, Cddl, model.parameters = list(p = list(formula = ~1),
                                                           c = list(formula = ~1)))

closedshared <- mark(data = Cproc, Cddl, model.parameters = 
                       list(p = list(formula=~1,share=TRUE)))

get.real(closed, "p", se = TRUE)
get.real(closed, "f0", se = TRUE)
get.real(closedshared, "p", se = TRUE)



pic1Dups_999 <- dup.fun(picnic_1[picnic_1$zoned == 1,], distance = 1, cluster = 10, TS = 9, TV = 9, TR = 9)

x999 <- pic1Dups_999[[1]] %>%
  mutate(dh = 1) %>%
  complete(transect_orientation, fixedDistDupID, 
           fill = list(dh=0)) %>%
  pivot_wider(id_cols = fixedDistDupID, values_from = dh, names_from = transect_orientation) %>%  
  unite(ch, EW:NS, sep="") 

x999 %>%
  filter(ch == "11")

## 66 observed by both

setwd("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScleroDistanceSampling")
x %>%
  mutate(ch = paste0(ch,";")) %>%
  write.table(file = "Data/Closed.inp",
              sep = " ", col.names = FALSE, row.names = FALSE)

scglClosed.inp <- convert.inp("Data/Closed.inp",
                              covariates = NULL,
                              use.comments = FALSE)

M <- nrow(x) ## 81 unique observations
clusterSize <- 1.58

## EW then NS
Cproc <- process.data(x, model = "Closed")
Cddl <- make.design.data(Cproc)

# constant p and c
closed <- mark(data = Cproc, Cddl, model.parameters = list(p = list(formula = ~1),
                                                           c = list(formula = ~1)))

closedshared <- mark(data = Cproc, Cddl, model.parameters = 
                       list(p = list(formula=~1,share=TRUE)))

get.real(closed, "p", se = TRUE)
get.real(closed, "f0", se = TRUE)
get.real(closedshared, "p", se = TRUE)
```

## MRDS  
```{r}
## Fixed distance threshold, select the nearest

pic1Dups_999 <- dup.fun(picnic_1[picnic_1$zoned == 1,], distance = 1, cluster = 10, TS = 9, TV = 9, TR = 9)

library(mrds)

picnic.data1 <- pic1Dups_999[[1]] %>%
  ## rename distance_to_group to distance, ignore "distance" which is the distance between observations
  dplyr::select(!distance) %>%
  rename(distance = distance_to_group) %>%
  mutate(detected = 1) %>%
  rename("object" = fixedDistDupID) %>%
  # rename("object" = fixedDistClustDupID) %>%
  rename("observer" = transect_orientation) %>%
  rowwise() %>%
  mutate(size = sum(c(number_seedlings_in_group, 
                      number_vegetative_in_group, 
                      number_reproductive_in_group), na.rm = TRUE)) %>%
  ungroup() %>%
  complete(nesting(object,size),observer, fill = list(detected = 0)) %>% ## Keep only size and plantID that are in the data
  dplyr::select(object,observer, size, detected, distance, transect_meter_mark) %>%
  arrange(object, observer) %>%
  mutate(observer = as.factor(as.numeric(as.factor(observer)))) %>%
  relocate(size, .after = distance) %>%
  mutate(object = as.integer(as.factor(object))) %>%
  ## When grouped, one might be missed, that's the row where distance missing
  group_by(object) %>%
  ## each set of two observations
  ## need to update the group size as well. 
  mutate(distance2 = case_when(detected == 0 & max(abs(transect_meter_mark), na.rm = TRUE) < 15 ~ abs(10-max(abs(transect_meter_mark), na.rm = TRUE)),
                        detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 15 & max(abs(transect_meter_mark), na.rm = TRUE) < 25 ~ 
                                                                                                 abs(20-max(abs(transect_meter_mark), na.rm = TRUE)),
                        detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 25 & max(abs(transect_meter_mark), na.rm = TRUE) < 35 ~ 
                                                                                                 abs(30-max(abs(transect_meter_mark), na.rm = TRUE)),
                        detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 35 & max(abs(transect_meter_mark), na.rm = TRUE) < 45 ~ 
                                                                                                 abs(40-max(abs(transect_meter_mark), na.rm = TRUE)),
                        detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 45 & max(abs(transect_meter_mark), na.rm = TRUE) < 55 ~ 
                                                                                                 abs(50-max(abs(transect_meter_mark), na.rm = TRUE)),
                        detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 55 & max(abs(transect_meter_mark), na.rm = TRUE) < 65 ~ 
                                                                                                 abs(60-max(abs(transect_meter_mark), na.rm = TRUE)),
      TRUE ~ max(distance, na.rm = TRUE)
                              )) %>%
  ungroup() %>%
  dplyr::select(object:detected,size,distance2) %>%
  rename("distance" = "distance2") %>%
  group_by(object,observer) %>%
  ## Where the size differs, only keep larger estimate of size; won't see more than are there, might miss some
  filter(size == max(size, na.rm = TRUE)) %>%
  ungroup() %>%
  distinct()
  

sizeVariation <- table(picnic.data1$object) ## A few have duplicate observations! so four instead of 2. Fixed by selecting larger
picnic.data1 %>%
  filter(object %in% c(sizeVariation != 2))

## Fully independent   
# MR only, no distance data
fi.mr.dist <- ddf(method = "trial.fi", 
                  mrmodel = ~ glm(link = 'logit', formula = ~ size), #  distance),
                  data = picnic.data1,
                  meta.data = list(width = 3))


detection.tables <- det.tables(fi.mr.dist)
# Print these detection tables
print(detection.tables)
plot(fi.mr.dist)

picnic.data1 %>%
  ggplot(  aes(distance, fill = observer)) +
    geom_histogram()


fi.mr.dist.io <- ddf(method = "io.fi", 
                  mrmodel = ~ glm(link = 'logit', formula = ~ size), #  distance),
                  data = picnic.data1,
                  meta.data = list(width = 3))


detection.tables.io <- det.tables(fi.mr.dist.io)
# Print these detection tables
print(detection.tables.io)
plot(fi.mr.dist.io)

############ Swap observer 1 and 2
picnic.data1.swapObs <- picnic.data1 %>%
  mutate(observer = case_when(observer == 1 ~ 2,
                              observer == 2 ~ 1))

fi.mr.dist.swap <- ddf(method = "trial.fi", 
                  mrmodel = ~ glm(link = 'logit', formula = ~ size), #  distance),
                  data = picnic.data1.swapObs,
                  meta.data = list(width = 3))


detection.tables.swap <- det.tables(fi.mr.dist.swap)
# Print these detection tables
print(detection.tables.swap)
plot(fi.mr.dist.swap)

```


2026.06.01 Michelle DePrenger-Levin additions from Hamilton et al. 2017   
inverse logit
```{r}
## Then repeat by swapping EW and NS? As trial - different form fully independent
## x == distance
## s == group size

# Probability of EW detection 
y_EW <- exp(alpha1 + beta1*x + gamma1*(x*x) + delta1*s + epsilon1*x*s) / 
          (exp(alpha1 + beta1*x + gamma1*(x*x) + delta1*s + epsilon1*x*s)+1)

# Probability of detection by NS conditional on non-detection of EW 
y_EW_notNS <-  exp(alpha2 + beta2*x + gamma2*(x*x) + delta2*s + epsilon2*x*s) / 
          (exp(alpha2 + beta2*x + gamma2*(x*x) + delta2*s + epsilon2*x*s)+1)

# Probability of detection by NS conditional on detection by EW (duplicates)
y_EW_NS <-  exp(alpha3 + beta3*x + gamma3*(x*x) + delta3*s + epsilon3*x*s) / 
          (exp(alpha3 + beta3*x + gamma3*(x*x) + delta3*s + epsilon3*x*s)+1)


```

