---
title: "2026_SCGL_work"
author: "Jonathan Amegin"
date: "2026-05-01"
output: html_document
editor_options: 
  markdown: 
    wrap: 90
---

Updates by Michelle DePrenger-Levin 2026.06.02

To do:

~~- for each observation, create clones in the data frame that switch the
"direction_from_transect" field for later duplicate analysis~~

~~- Split the comparison of group sizes up between seedlings, vegetative, and
reproductive~~

~~-   Develop probability scoring to show 'percent similarity' between observations based on
    weighted scores of covariates

~~-   Keep only the duplicates with the shortest distance between the compared observations
    so that only 1 duplicate/global_id

##Setting Up

```{r message=FALSE, warning=FALSE, include=FALSE}
# tidyverse contains both 'dyplr' and 'ggplot2'
library(tidyverse)
library(janitor)
library(uuid)

setwd("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Sclerocactus-glaucus/Sclerocactus-glaucus_RawData/2024_CrossedDistanceSampling")
#this contains all observations of clusters of SCGL for all sites & Transects
scgl_obs <- read_csv("distanceSampling_1.csv")
scgl_obs <- clean_names(scgl_obs)

#This contains metadata for each transect including length & direction
scgl_trans_length <- read_csv("transect_2.csv")
scgl_trans_length <- clean_names(scgl_trans_length)

#filtering NA's and outliers, and changing some variable classes for later work
scgl <- scgl_obs[!(is.na(scgl_obs$transect_meter_mark|scgl_obs$distance_to_group)),] %>%  
        filter(transect_meter_mark < 110, distance_to_group < 10) %>% 
        mutate( object_id = as.factor(object_id),
                site_name = as.factor(site_name),
                transect_number = as.factor(transect_number),
                swapped_dir = 0)

scgl[is.na(scgl[,"number_seedlings_in_group"]), "number_seedlings_in_group"] <- 0
scgl[is.na(scgl[,"number_vegetative_in_group"]), "number_vegetative_in_group"] <- 0
scgl[is.na(scgl[,"number_reproductive_in_group"]), "number_reproductive_in_group"] <- 0

#adding clone observations with the 'direction_to_group' swapped for later analysis
#some observations had NA, so need to check both directions for them
### The NA+ and NA-  see lines 74, default is + for direction
scgl1 <- rbind(scgl,
               mutate(scgl,
                      swapped_dir = "1",
                      direction_to_group = case_when(direction_to_group == "+" ~ "-",
                                                     direction_to_group == "-" ~ "+",
                                                     is.na(direction_to_group) ~ "NA-"
                                                     )))

scgl1[is.na(scgl1[,"direction_to_group"]), "direction_to_group"] <- "NA+"

#use the attributes to add modifiers to the measurements
#EW transect meter marks need to be * -1 because they were recorded as if in the(+,+) quadrant, but the mapping we will do is based on (-,+) quadrant.             
scgl1 <-
  scgl1 %>%
  mutate(
    transect_meter_mark = case_when(
      transect_orientation == "EW" ~ transect_meter_mark * -1,
      TRUE ~ transect_meter_mark),
    #making the 'direction to group' indicate if the 'distance to group' needs to be left/right of transect.  
    distance_to_group = case_when(
      direction_to_group == "-" | direction_to_group == "NA-" ~ distance_to_group * -1,
                                                         TRUE ~ distance_to_group
      ))
           
#setting up the (X,Y) columns to then transform based on plot/shapefile analysis
scgl1 <- scgl1 %>% 
  mutate( x = case_when(transect_orientation == "NS" ~ distance_to_group,
                        transect_orientation == "EW" ~ transect_meter_mark), 
          y = case_when(transect_orientation == "EW" ~ distance_to_group,
                        transect_orientation == "NS" ~ transect_meter_mark)) 

#joining the two dataframes to add transect metadata to each observation
scgl1 <- left_join(scgl1, scgl_trans_length[, c("transect_length", "global_id")],
                   by = join_by(parent_global_id == global_id))                
                
#filtering to sites
fram <- scgl1 %>% 
  filter(site_name == "12")

picnic <- scgl1 %>% 
  filter(site_name == "1"| site_name == "3")

bridgeport <- scgl1 %>% 
  filter(site_name == "10")

oilpad <- scgl1 %>% 
  filter(site_name == "11")

```



##picnic
```{r}
#transforming coordinates to match the plot design. This requires accurate note taking
 picnic_1 <-  picnic %>% 
   mutate( 
    x = case_when(transect_orientation == "NS" & transect_number == "1" ~ x - 20, 
                  transect_orientation == "NS" & transect_number == "2" ~ x - 30,
                  transect_orientation == "NS" & transect_number == "3" ~ x - 40,
                  transect_orientation == "EW" & transect_number == "1" ~ x - 20,
                  transect_orientation == "EW" & transect_number == "2" ~ x - 10,
                  transect_orientation == "EW" & transect_number == "3" ~ x - 10,
                  transect_orientation == "EW" & transect_number == "4" ~ x
                  ),
    y = case_when(transect_orientation == "NS" & transect_number == "1" ~ y + 20, 
                  transect_orientation == "NS" & transect_number == "2" ~ y + 10,
                  transect_orientation == "NS" & transect_number == "3" ~ y ,
                  transect_orientation == "EW" & transect_number == "1" ~ y + 20,
                  transect_orientation == "EW" & transect_number == "2" ~ y + 30,
                  transect_orientation == "EW" & transect_number == "3" ~ y + 40,
                  transect_orientation == "EW" & transect_number == "4" ~ y + 50,
                  ))

#adding an attribute to show if an observation is within the zone that NS & EW transects share
picnic_1 <-
  mutate(picnic_1,
         zoned = case_when(
           c(rowSums(mapply(FUN = function(a, b){between(a, b-3, b+3)}, picnic_1[, "y"], c(20, 30, 40, 50))) +
             rowSums(mapply(FUN = function(a, b){between(a, b-3, b+3)}, picnic_1[, "x"], c(-20, -30, -40)))) == 2 ~ 1,
           TRUE ~ 0))

## Add unique ID for each zone, only compare plants within a zone, not across all plants


```

### 2026 function

Idea is to use covariates to generate a probablity score indicating likelihood of pairs of
observations being duplicates.

Covariates affecting probability:

-   Number in groups, split between categories.

-   Distance between observations

-   Direction of transect

-   within/out the crossed area

function to apply comparison to each object_id pair by referencing the fram_1 table Needs
to use the combn() pairs and parse the fram_1 table and then compare the covariates to
evaluate a probability score
```{r}
## testing
site <- picnic_1[picnic_1$zoned == 1,]
Tdistance <- 1
# Tcluster <- 12
# TS <- 5
# TV <- 2
# TR <- 1

## doesn't also have the reverse
# foo <- utils::combn(1:3, m = 2, simplify = FALSE)
# do.call(rbind,foo)
# ## expand.grid does give both 1 2 and 2 1
# foo1 <- expand.grid(x = 1:3, y = 1:3)
# foo1

## Works to get down to one or the other swapped for each plant
 # foo <- data.frame(objectID = sample(1:10, 100, replace = TRUE),
 #                   swapped = rbinom(100,1,prob = 0.1))
 # 
 # foo %>%
 #   group_by(objectID, swapped) %>%
 #   summarise(n = n())
 # 
 # foo2 <- foo %>%
 #   group_by(objectID) %>%
 #   arrange(desc(swapped)) %>%
 #   filter(row_number() == 1) %>%
 #   ungroup() %>%
 #   arrange(objectID)

data.frame(do.call(rbind,list)) %>%
  group_by(X1 == X8, X10) %>% # SwapMe for a, the first, the transect of the first
  summarise(n = n())  ### 11 times the first was swapped, all from NS

data.frame(do.call(rbind,list)) %>%
  group_by(X2 == X9, X11) %>% # SwapMe for a, the first
  summarise(n = n())  ### 4 times the second was swapped, all from NS

## if 10 are duplicate observations, then 81 groups, 20 from NS, 71 from EW
site %>%
  filter(swapped_dir == 0 &
           zoned == 1) %>%
  group_by(transect_orientation) %>%
  summarise(n = n())

## Already only the zoned ones; still has all the swapped, need to remove
siteout %>%
  group_by(transect_orientation, swapped_dir) %>%
  summarise(n = n())
```


```{r}
###Custom Function 1
# Input plot data mapped to (X, Y) grid and plants added with the direction_to_group swapped - test uncertainty of direction
dup.fun <- function(site, Tdistance = 1){ # , Tcluster = 12, TS = 8, TV = 2, TR = 2){
  
  #making new column with new UUID for each observation due to cloned entries
  site$new_id = UUIDgenerate(use.time = FALSE, n = nrow(site), output = "string")
  
  #Creating list of all combinations of global_id's to reference, order doesn't matter, start with only original
  list <- utils::combn(site$new_id[site$swapped_dir == 0], m = 2, simplify = FALSE)

  # (1) first go through and keep or discard a swapped plant placement,
  ## But, doesn't matter if they're on the same transect, doesn't matter if they are far away, Only getting duplicate, use that to 
  #Appending calculated variables to the list  
  list <- lapply(list, function(list.1){

    #indexing data with the paired global_ids
    a <- site[site$new_id == list.1[1],]
    b <- site[site$new_id == list.1[2],]

    # Check the swapped 
    a1 <- site[site$object_id == a$object_id & site$swapped_dir == 1,]
    b1 <- site[site$object_id == b$object_id & site$swapped_dir == 1,]

    #Calculating distance, if the swapped is closer, then always use the swapped, only if on different transects
    # do both for all, then only keep the original or swapped if_any are marked that the swapped is closer
    ## Need a case_when for both
    #  They could be a duplicate if (1) they are from different transect orientations
    #                               (2) the pair (or any combination of swapped directions) are closer than the threshold distance 
    ### In this case "a" and "b" represent the original mapped position, "a1" and "b1" are the swapped direction
    # if(a$transect_orientation != b$transect_orientation){
      Original <- dist(rbind(c(a$x,a$y), c(b$x,b$y)), method = "euclidean")
      SwappedA <- dist(rbind(c(a1$x,a$y), c(b$x,b$y)), method = "euclidean")
      SwappedB <- dist(rbind(c(a$x,a$y), c(b1$x,b$y)), method = "euclidean")
      SwappedAB <- dist(rbind(c(a1$x,a$y), c(b1$x,b$y)), method = "euclidean")
      
      if(any(c(Original,SwappedA,SwappedB,SwappedAB) < Tdistance)){
        ## Which should be swapped
        SwapMe <- c("a_b","a1_b","a_b1","a1_b1")[which.min(c(Original,SwappedA,SwappedB,SwappedAB))]
        list.1[3] <- min(Original, SwappedA, SwappedB, SwappedAB)
      } else {
        list.1[3] <- dist(rbind(c(a$x,a$y), c(b$x,b$y)), method = "euclidean")
        SwapMe <- "a_b"
      }
      
    # } else {
    #   list.1[3] <- dist(rbind(c(a$x,a$y), c(b$x,b$y)), method = "euclidean")
    #   SwapMe <- "a_b"
    # }
                           

    #determining if they are on the same transect
    ## same == 1, different == 0
    list.1[4] <- case_when(a$transect_orientation == b$transect_orientation ~ 1,
                           a$transect_orientation != b$transect_orientation ~ 0)

    
    #Comparing if the 'direction_from_transect' variable was swapped earlier
    list.1[5] <- case_when(any(sapply(SwapMe, str_detect, "a1")) ~ a1$new_id,
                           TRUE ~ a$new_id) ## id_1 updated
    list.1[6] <- case_when(any(sapply(SwapMe, str_detect, "b1")) ~ b1$new_id,
                           TRUE ~ b$new_id) ## id_2 updated
    #The orientation of id_1; id_2
    list.1[7] <- a$transect_orientation
    list.1[8] <- b$transect_orientation
    
    # First or second swapped? 
    list.1[9] <- case_when(any(sapply(SwapMe, str_detect, "a1")) ~ 1,
                           TRUE ~ 0)
    list.1[10] <- case_when(any(sapply(SwapMe, str_detect, "b1")) ~ 1,
                           TRUE ~ 0) ## id_1_updated
    
    list.1[11] <- a$object_id
    list.1[12] <- b$object_id
    
     return(list.1)
  })
  

    df <- data.frame(do.call(rbind,list))
    names(df) <- c("V1_id1", "V2_id2",
                      "distanceBetween",
                      "same_diff",  ## if on same or different transect 1 = same, 0 = different
                      "V1_updatedid",  # after checking if need to be swapped
                      "V2_updatedid",  # after checking if need to be swapped
                      "V1_orientation",
                      "V2_orientation",
                      "V1_swapped",
                      "V2_swapped",
                      "V1_objectID",
                      "V2_objectID")

 df$dupID <- UUIDgenerate(use.time = FALSE, n = nrow(df), output = "string") 

 # ## Wide to long isn't working correctly
 # df1 <- df %>%
 #   filter(same_diff == 0) %>%
 #   filter(distanceBetween < Tdistance) %>%
 #   select(V1_id1, distanceBetween, V1_updatedid, V1_orientation, V1_swapped, V1_objectID) %>%
 #   group_by(V1_id1) %>%
 #   filter(distanceBetween == min(distanceBetween))
 # df2 <- df %>%
 #   filter(same_diff == 0) %>%
 #   filter(distanceBetween < Tdistance) %>%
 #   select(V2_id2, distanceBetween, V2_updatedid, V2_orientation, V2_swapped, V2_objectID) %>%
 #   group_by(V2_id2) %>%
 #   filter(distanceBetween == min(distanceBetween))
 ## df wide to long splitting transect orientation and id; keep only the swapped or original. 
 df_long <- df %>%
   filter(same_diff == 0) %>%
   filter(distanceBetween < Tdistance) %>%
   pivot_longer(cols = starts_with("V"), names_to = c("ID", ".value"),
                names_sep = "_") %>%
   group_by(objectID) %>%
   filter(distanceBetween == min(distanceBetween)) %>%
   arrange(desc(swapped)) %>%
   filter(row_number() == 1) %>%
   ungroup() %>%
   select(!c(id1, id2, ID, same_diff)) %>%
   mutate(objectID = as.numeric(objectID))

 # The updated ID is the swapped or original (if swapped is < Tdistance) 
 # the dupID should be applied to the updatedid when < Tdistance
 siteout <- site %>%
   mutate(object_id = as.numeric(object_id)) %>%
   left_join(df_long, by = c("new_id" = "updatedid")) %>%
   mutate(dupID = case_when(is.na(dupID) ~ new_id,
                            TRUE ~ dupID)) %>%
   filter(case_when(swapped == 0 ~ swapped_dir == 0,
                    swapped == 1 ~ swapped_dir == 1))
   # group_by(object_id) %>%
   # filter(distanceBetween == min(distanceBetween))

 return(siteout)
}

```

```{r}

siteout %>%
  select(object_id, distance_to_group, objectID)
  
# objectID should match object_id once merged
df_long %>%
  group_by(objectID) %>%
  summarise(n = n()) %>%
  filter(n != 1)

## Should be only duplicate observations, means there are 14 duplicates when you swap the direction from the transect of 8 of the NS observations. But there were 10 before when you didn't swap sides of any. 
df_long %>%
  group_by(swapped, orientation) %>%
  summarise(n = n())

## Only 6 swapped?
siteout %>%
  group_by(swapped, transect_orientation) %>%
  summarise(n = n())

## 13 duplicate observations
siteout %>%
  group_by(dupID) %>%
  summarise(n = n()) %>%
  filter(n > 1) ## 13 duplicates

# do we only have either the swapped or the original of each object ID?
## Yes, probably because only in some comparisons is the swapped picked -- fixed with distanceBetween == min()
# Now no!
siteout %>%
  group_by(object_id) %>%
  summarise(n = n()) %>%
  filter(n > 1)

## Why is objectID different?
siteout %>%
  filter(object_id == 109) %>%
  select(object_id, distance_to_group, swapped_dir, distanceBetween, dupID, id1, id2, objectID)

## How often did it swap? Not very often!
data.frame(do.call(rbind, list)) %>%
  mutate(SwappedAorB = case_when((X1 != X8 & X2 == X9) ~ "a1",
                                 (X2 != X9 & X1 == X8) ~ "b1",
                                 (X1 != X8 & X2 != X9) ~ "a1b1",
                                 TRUE ~ "original")) %>%
  group_by(SwappedAorB) %>%
  summarise(n = n())



data.frame(do.call(rbind, list)) %>%
  filter(is.na(X3))

picnic_1 %>%
  filter(swapped_dir == 0) %>%
  ggplot( aes(x,y, color = transect_orientation)) +
    geom_point()

```


```{r}
####Custom Function 2
#checking for observations that have multiple duplicates & then filtering down to 'most likely'
## If discrepancy is over 1, then for sure isn't a duplicate! Filters down to the ones that 'could' be 
## fixed_tolerance == 1: cluster size and distance is under tolerance, 0: over tolerance
## add a fixed_tolerance for just distance
dup.fun.1 <-
  function(site){
  df <- filter(site, discrepency <1 & orientation == 0) #filter to possible dupes, removed known not duplicates
  id <- c(df$id_1, df$id_2) #vector of all id's that are associated with duplicated observations
  if (sum(duplicated(id)) != 0){
    dup.id <- unique(id[duplicated(id)])
    id.1 <- #function to find the multiples that are NOT the most likely
    lapply(dup.id, function(i){
        df.1 <- filter(df, id_1 == i | id_2 == i)
        df.2 <- filter(df.1, discrepency != min(discrepency))$dup_id
        df.2
    })
    df[-which(df$dup_id %in% unlist(id.1)),]}
  else{df} 
}

###Custom Function 3
# Getting the numbers needed for the unmakred package: Unmakred dulicate observation function
# This function calculates the number of observations made by each observer (different transect directions) 
# alongside the number of observations said to be duplicates based on the earlier dup.fun function
# and outputs a named list
## Filters to the minimum discrepancy score

# Use only original or also swapped: swap is true and false (use swapped or not)
# 
umdo.fun <- 
function(site, swap, d, c){
  #allows choice of using the cloned/mirrored observations for checking errors
  if(swap == FALSE){ 
  site.1 <- filter(site, swapped_dir == 0 & zoned == 1) # only the zoned ones
  site.ew <- filter(site.1, transect_orientation == "EW")
  site.ns <- filter(site.1, transect_orientation == "NS")
  site.1d <- dup.fun(site.1, distance = d, cluster = c)
  #getting the numbers of duplicates for both the discrepancy score & fixed tolerance 
  #analysis allows for comparison between methods
  site.score <- dup.fun.1(filter(site.1d, discrepency < 1))
  site.fixed <- dup.fun.1(filter(site.1d, fixed_tolerance == 1))
  # site.fdistance <- 
  
  
  list <- list("EW" = nrow(site.ew), 
               "NS" = nrow(site.ns),
               "scored" = nrow(site.score),
               "fixed" = nrow(site.fixed))
  }else{
  site.1 <- filter(site, zoned == 1)
  site.ew <- filter(site.1, transect_orientation == "EW" & swapped_dir == 0)
  site.ns <- filter(site.1, transect_orientation == "NS" & swapped_dir == 0)
  site.1d <- dup.fun(site.1, distance = d, cluster = c)
  #getting the numbers of duplicates for both the discrepency score & fixed tolerance 
  #analysis allows for comparison between methods
  site.score <-   dup.fun.1(filter(site.1d, discrepency < 1 & swapped_dir == 0)) # neither swapped
  site.score.1 <- dup.fun.1(filter(site.1d, discrepency < 1 & swapped_dir < 2)) # one swapped
  site.score.2 <- dup.fun.1(filter(site.1d, discrepency < 1 & swapped_dir <= 2)) # two or one swapped
  site.fixed <-   dup.fun.1(filter(site.1d, fixed_tolerance == 1 & swapped_dir == 0))
  site.fixed.1 <- dup.fun.1(filter(site.1d, fixed_tolerance == 1 & swapped_dir < 2))
  site.fixed.2 <- dup.fun.1(filter(site.1d, fixed_tolerance == 1 & swapped_dir <= 2))
  ## and fixed distance
  
  
  list <- list("EW" = nrow(site.ew), 
               "NS" = nrow(site.ns),
               "scored" = nrow(site.score),
               "scored swap 1" = nrow(site.score.1),
               "scored swap 2" = nrow(site.score.2),
               "fixed" = nrow(site.fixed),
               "fixed swap 1" = nrow(site.fixed.1),
               "fixed swap 2" = nrow(site.fixed.2))
  }
  return(list)
}


```



##Fram

###Plotting individuals

Transforming Fram coordinates to accurately reflect plot design.This is based on the
origin (0,0) being in the lower right of the plot so that one or both of the axis matches
the start of at least one transect. In this case, the origins of the three NS transects
are at Y = 0, and 4 out of 5 EW transects' origins are at X = 0.

```{r fram transform}

fram_1 <- fram %>% 
  mutate( 
    x = case_when(transect_orientation == "NS" & transect_number == "1" ~ x - 10, 
                  transect_orientation == "NS" & transect_number == "2" ~ x - 20,
                  transect_orientation == "NS" & transect_number == "3" ~ x - 30,
                  transect_orientation == "EW" & transect_number == "5" ~ x - 10,
                  TRUE ~ x
                  ),
    y = case_when(transect_orientation == "EW" & transect_number == "1" ~ y + 10,
                  transect_orientation == "EW" & transect_number == "5" ~ y + 50,
                  transect_orientation == "EW" & transect_number == "7" ~ y + 70,
                  transect_orientation == "EW" & transect_number == "8" ~ y + 80,
                  transect_orientation == "EW" & transect_number == "9" ~ y + 90,
                  TRUE ~ y),
    )

#Fram NS lines tapes were laid backwards making 0 = 100. 
fram_1 <- fram_1 %>% 
  mutate( y = case_when(transect_orientation =="NS" ~ ((y - 100)* -1), TRUE ~ y)
    )

fram_1 <- mutate(fram_1, zoned = 
                   case_when(
                     c(rowSums(mapply( FUN = function(a, b){between(a, b-3, b+3)}, fram_1[, "y"], c(10, 50, 70, 80, 90))) +
                                                rowSums(mapply( FUN = function(a, b){between(a, b-3, b+3)}, fram_1[, "x"], c(-10, -20, -30)))) == 2 ~ 1,
       TRUE ~ 0)) 
```

```{r fram graphing}
#transect start/stops for mapping
fram_lines <- tibble( x = c(-10, -20, -30, 0, -10, 0, 0, 0), 
                      y = c(0, 0, 0, 10, 50, 70, 80, 90), 
                      xend = c(-10, -20, -30, -33, -33, -60, -65, -39), 
                      yend = c(100, 100, 100, 10, 50, 70, 80, 90),
                      label = c("1NS", "2NS", "3NS", "1EW", "5EW", "7EW", "8EW", "9EW")
                    )


 fram_plot<-  ggplot( data = fram_1, mapping = aes(x = x, y = y)) +
  geom_point(size = 0.3) +
  geom_segment(data = fram_lines, aes(x = x, y = y, xend = xend, yend = yend)) +
   geom_text(data = fram_lines,
            aes( label = label),
            nudge_x = 4,
            size = 2,
            )

  # should expect to see mirroring of each point, originals & mirrors are colored 
  fram_plot +
    geom_point(data = filter(fram_1, zoned ==1), aes( colour = transect_orientation))+
    coord_fixed(ratio = 1, xlim = c(-75, 10), ylim = c(0, 100))
   # geom_line(aes( group = global_id), color = "red")

  fram_plot
  
  
fram_1 %>%
  dplyr::select(x,y,swapped_dir,object_id, transect_orientation)


  
  fram_1 %>%
    # filter(swapped_dir == 0) %>%
    mutate(swapped_dir = as.factor(swapped_dir)) %>%
ggplot(  aes(x,y) ) +
  geom_point(data = fram_1,  aes(x,y, shape = swapped_dir, 
                                 color = transect_orientation),
             size = 4) +
  geom_segment(data = fram_lines, aes(x = x, y = y, xend = xend, yend = yend)) +
   # geom_text(data = fram_lines, 
   #          aes( label = label),
   #          nudge_x = 4,
   #          size = 2,
   #          ) +
    scale_shape_manual(values = c(16,1)) +
    xlim(c(-16.9,-23.1)) +
    ylim(c(46.9,53.1)) +
  geom_text(data = fram_1, aes(label = number_reproductive_in_group),
            size = 2) +
    coord_fixed(ratio = 1) +
    scale_colour_manual(values = c("yellow","skyblue")) +
  geom_text(data = fram_1, aes(label = object_id),
            nudge_x = 0.25,
            nudge_y = 0.25,
            size = 2.5) +
    coord_fixed(ratio = 1) 
  
fram_1 %>%
  filter(x > -23.1 & x < -16.9 &
           y > 46.9 & y < 53.1) %>%
  bind_rows(fram_1) %>%
  mutate()
  dplyr::select(x, y, object_id, transect_orientation, swapped_dir, number_reproductive_in_group, number_vegetative_in_group) %>%
  arrange(x, y)
  
```

Testing
```{r}

site <- fram_1
#making new column with new UUID for each observation due to cloned entries
  site$new_id = UUIDgenerate(use.time = FALSE, n = nrow(site), output = "string")
  #Creating list of all combinations of global_id's to reference
  list <- utils::combn(site$new_id, m = 2, simplify = FALSE)
  #Appending calculated variables to the list  
  list <- lapply(list, function(list.1){

    #indexing data with the paired global_id's
    a <- site[site$new_id == list.1[1],]
    b <- site[site$new_id == list.1[2],]

    #Calculating distance
    list.1[3] <- dist(rbind(c(a$x,a$y), c(b$x,b$y)), method = "euclidean")

    #determining if they are on the same transect
    list.1[4] <- case_when(a$transect_orientation == b$transect_orientation ~ 1,
                           a$transect_orientation != b$transect_orientation ~ 0)

    #Group size comparisons
    list.1[5] <- abs(diff(c(a$number_seedlings_in_group,b$number_seedlings_in_group)))
    list.1[6] <- abs(diff(c(a$number_vegetative_in_group,b$number_vegetative_in_group)))
    list.1[7] <- abs(diff(c(a$number_reproductive_in_group,b$number_reproductive_in_group)))
    
    #Comparing if the 'direction_from_transect' variable was swapped earlier
    #attempts to correct for potential recorder error
    #could change to a calculation to make weighting the probability score easier
    list.1[8] <- as.numeric(a$swapped_dir) + as.numeric(b$swapped_dir)
    list.1[9] <- case_when(str_detect(a$direction_to_group, pattern = "NA") &
                            str_detect(b$direction_to_group, pattern = "NA") ~ 2,
                           xor(str_detect(a$direction_to_group, pattern = "NA"), 
                            str_detect(b$direction_to_group, pattern = "NA")) ~ 1,
                            TRUE ~ 0)
     return(list.1)
  })
  
all.fram <- do.call(rbind, list) %>%
  as.data.frame() %>%
  filter(V3 < 1)
```



Looks like EW 291 matches NS 340 when EW 291 IS NOT swapped but NS 340 IS swapped     
Looks like EW 290 matches NS 


```{r}
#This is just testing some things out, safe to ignore 
test.fram <- dup.fun(filter(fram_1, zoned == 1), distance = 0.25, cluster = 5)

filter(test.fram, swapped_dir == 0)$discrepency %>% 
  hist(labels = TRUE)

nrow(filter(test.fram, discrepency < 0.5))

sapply(seq(0,1,0.05), function(i){
  nrow(filter(test.fram,  swapped_dir == 0 & discrepency < i))
})

umdo.fun(fram_1, swap = FALSE)
umdo.fun(fram_1, swap = TRUE)


  # filter(test.fram, distance <8, discrepency <1)$distance %>% 
  # hist( breaks = 30)
```


```{r}
# umarked.picnic <- umdo.fun(picnic_1, swap = TRUE, d = 1, c = 5)

# [[1]]: the input site with the unique ID (new_id) added and the duplicates identified according to fixed distance, fixed distance and cluster size cutoffs, and rbinom for discrepency scores under 1 (over 1 considered not a duplicate)
# [[2]]: the data frame of paired comparisons
# [[3]]: the data frame in long form with unique values (instead of comparison of 1 vs 2 and 2 vs 1 in the wide data frame)
pic1Dups <- dup.fun(picnic_1[picnic_1$zoned == 1,], distance = 1, cluster = 4, TS = 8, TV = 3, TR = 2)
pic1Dups_821 <- dup.fun(picnic_1[picnic_1$zoned == 1,], distance = 0.5, cluster = 10, TS = 8, TV = 2, TR = 1)
## Higher discrepancy score == less likely to be a duplicate. Lower discrepancy score (approaching zero) == more likely duplicate
# #4 "orientation" 1 == same transect, 0 == different transect orientation
# #5 "diff_seedlings" diff seedlings; #6 diff veg; #7 diff repro
# #10 is a sum of how much distance is less than threshold +
#                 how much seedling + veg + rep is less than threshold

# Only slightly lower discrepancy score for those on different grid lines. Not surprising 
### 1 == same transect (known NOT duplicates)
### 0 == different transects (possible duplicates)
pic1Dups[[2]] %>% 
  filter(fixed_tolerance == 1) %>%
  ggplot(  aes(discrepency, fill = as.factor(same_diff)))+
    geom_density(alpha = 0.5) +
    theme_bw() 

pic1Dups[[2]] %>% 
  filter(fixed_disttolerance == 1) %>%
  ggplot(  aes(discrepency, fill = as.factor(same_diff)))+
    geom_density(alpha = 0.5) +
    theme_bw() 

# distance and cluster size
pic1Dups_821[[2]] %>% 
  filter(fixed_tolerance == 1) %>%
  ggplot(  aes(discrepency, fill = as.factor(same_diff)))+
    geom_density(alpha = 0.5) +
    theme_bw()

pic1Dups[[2]] %>%
  rowwise() %>%
  mutate(discrepYN = rbinom(1,1, prob = discrepency)) %>%
  filter(discrepYN == 0) %>%
  ggplot(  aes(discrepency, fill = as.factor(same_diff)))+
    geom_density(alpha = 0.5) +
    theme_bw()

pic1Dups[[2]] %>%
  filter(discrepency < 1)
## 
pic1Dups[[2]] %>%
  filter(discrepency < 1) %>%
  rowwise() %>%
  mutate(discrepYN = rbinom(1,1, prob = discrepency)) %>%
  # filter(same_diff == 0) %>%
  ggplot(  aes(distance, discrepYN, color = as.factor(same_diff))) +
    geom_point(pch = "|")+
    stat_smooth(method="glm",  method.args = list(family = "binomial"), se=TRUE) +
    theme_bw() +
    scale_color_discrete("Transect", labels = c("Possible duplicates","Known NON-duplicates"))

pic1Dups[[2]] %>%
  filter(discrepency < 1) %>%
  rowwise() %>%
  mutate(discrepYN = rbinom(1,1, prob = discrepency)) %>%
  # filter(same_diff == 0) %>%
  ggplot(  aes(distance, discrepYN, color = as.factor(same_diff))) +
    geom_point(pch = "|")+
    stat_smooth(method="glm",  method.args = list(family = "binomial"), se=TRUE) +
    theme_bw() +
    scale_color_discrete("Transect", labels = c("Possible duplicates","Known NON-duplicates"))

## Zero are the duplicates
pic1Dups[[2]] %>%
  filter(discrepency < 1) %>%
  rowwise() %>%
  mutate(discrepYN = rbinom(1,1, prob = discrepency)) %>%
  filter(same_diff == 0) %>%
  ggplot(  aes(distance, diff_reproductive)) +
    geom_point()+
    theme_bw() +
    facet_wrap(~ discrepYN)


```

ids are duplicated in picnic_1 for the original and the swapped  
## Need to remove all the swapped ones that are not closer. 
group_by plant Id and only keep the one with the closer distance (if within the threshold) or only keep original.  
    case_when pair with other transect is closer and < 1 meter ~ keep this one
    case_when pair has neither swapped or original within 1 meter, keep original
    case_when direction_to_group was NA, keep the one with a closer distance to another one?? | if none within 1 meter, keep one at random. 
```{r}
picnic_1 %>%
  ggplot(  aes(object_id)) +
    geom_histogram(stat = "count") +
    facet_wrap(~ swapped_dir)
```


Mark-Recapture   
fixedDistDupID
```{r}
library(RMark)

## Should be 91 records out
picnic_1 %>%
  filter(zoned == 1 & swapped_dir == 0)

# only zoned
pic1Dups <- dup.fun(picnic_1[picnic_1$zoned == 1,], Tdistance = 1)

## Not only 32 plants!!! where's the other ones?
x <- pic1Dups %>%
  mutate(dh = 1) %>%
  complete(transect_orientation, dupID, 
           fill = list(dh=0)) %>%
  pivot_wider(id_cols = dupID, values_from = dh, names_from = transect_orientation) %>%  
  unite(ch, EW:NS, sep="") 

x %>%
  filter(ch == "11")

## 13 observed by both

setwd("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScleroDistanceSampling")
x %>%
  mutate(ch = paste0(ch,";")) %>%
  write.table(file = "Data/Closed.inp",
              sep = " ", col.names = FALSE, row.names = FALSE)

scglClosed.inp <- convert.inp("Data/Closed.inp",
                              covariates = NULL,
                              use.comments = FALSE)

M <- nrow(x) ## 81 unique observations
clusterSize <- 1.58

## EW then NS
Cproc <- process.data(x, model = "Closed")
Cddl <- make.design.data(Cproc)

# constant p and c
closed <- mark(data = Cproc, Cddl, model.parameters = list(p = list(formula = ~1),
                                                           c = list(formula = ~1)))

closedshared <- mark(data = Cproc, Cddl, model.parameters = 
                       list(p = list(formula=~1,share=TRUE)))

get.real(closed, "p", se = TRUE)
get.real(closed, "f0", se = TRUE)
get.real(closedshared, "p", se = TRUE)




```

## MRDS  
```{r}

library(mrds)

picnic.data1 <- pic1Dups %>%
  ## rename distance_to_group to distance
  rename(distance = distance_to_group) %>%
  mutate(detected = 1) %>%
  rename("object" = dupID) %>%
  rename("observer" = transect_orientation) %>%
  rowwise() %>%
  mutate(size = sum(c(number_seedlings_in_group, 
                      number_vegetative_in_group, 
                      number_reproductive_in_group), na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(object) %>%
  mutate(size = max(size, na.rm = TRUE)) %>%
  ungroup() %>%
  complete(nesting(object,size),observer, fill = list(detected = 0)) %>% ## Keep only size and plant ID that are in the data
  dplyr::select(object,observer, size, detected, distance, transect_meter_mark) %>%
  arrange(object, observer) %>%
  mutate(observer = as.factor(as.numeric(as.factor(observer)))) %>%
  relocate(size, .after = distance) %>%
  mutate(object = as.integer(as.factor(object))) %>%
  ## When grouped, one might be missed, that's the row where distance missing
  group_by(object) %>%
  ## each set of two observations
  ## need to update the group size as well. 
  mutate(distance2 = case_when(
    detected == 0 & max(abs(transect_meter_mark), na.rm = TRUE) < 15 ~ abs(10-max(abs(transect_meter_mark), na.rm = TRUE)),
    detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 15 & max(abs(transect_meter_mark), na.rm = TRUE) < 25 ~ 
                                                                             abs(20-max(abs(transect_meter_mark), na.rm = TRUE)),
    detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 25 & max(abs(transect_meter_mark), na.rm = TRUE) < 35 ~ 
                                                                             abs(30-max(abs(transect_meter_mark), na.rm = TRUE)),
    detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 35 & max(abs(transect_meter_mark), na.rm = TRUE) < 45 ~ 
                                                                             abs(40-max(abs(transect_meter_mark), na.rm = TRUE)),
    detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 45 & max(abs(transect_meter_mark), na.rm = TRUE) < 55 ~ 
                                                                             abs(50-max(abs(transect_meter_mark), na.rm = TRUE)),
    detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 55 & max(abs(transect_meter_mark), na.rm = TRUE) < 65 ~ 
                                                                             abs(60-max(abs(transect_meter_mark), na.rm = TRUE)),
      TRUE ~ max(distance, na.rm = TRUE)
                              )) %>%
  ungroup() %>%
  dplyr::select(object:detected,size,distance2) %>%
  rename("distance" = "distance2") %>%
  group_by(object,observer) %>%
  ## Where the size differs, only keep larger estimate of size; won't see more than are there, might miss some
  filter(size == max(size, na.rm = TRUE)) %>%
  ungroup() %>%
  distinct()
  
picnic.data1 %>%
  group_by(observer, detected) %>%
  summarise(n = n())

picnic.data1 %>%
  group_by(object) %>%
  filter(detected == 1) %>%
  summarise(n = n()) %>%
  filter(n > 1)

sizeVariation <- table(picnic.data1$object) ## A few have duplicate observations! so four instead of 2. Fixed by selecting larger
picnic.data1 %>%
  filter(object %in% c(sizeVariation != 2))

## Fully independent   
# MR only, no distance data
fi.mr.dist <- ddf(method = "trial.fi", 
                  mrmodel = ~ glm(link = 'logit', formula = ~ size), #  distance),
                  data = picnic.data1,
                  meta.data = list(width = 3))


detection.tables <- det.tables(fi.mr.dist)
# Print these detection tables
print(detection.tables)
plot(fi.mr.dist)

picnic.data1 %>%
  ggplot(  aes(distance, fill = observer)) +
    geom_histogram()


fi.mr.dist.io <- ddf(method = "io.fi", 
                  mrmodel = ~ glm(link = 'logit', formula = ~ size), #  distance),
                  data = picnic.data1,
                  meta.data = list(width = 3))


detection.tables.io <- det.tables(fi.mr.dist.io)
# Print these detection tables
print(detection.tables.io)
plot(fi.mr.dist.io)

############ Swap observer 1 and 2
picnic.data1.swapObs <- picnic.data1 %>%
  mutate(observer = case_when(observer == 1 ~ 2,
                              observer == 2 ~ 1))

fi.mr.dist.swap <- ddf(method = "trial.fi", 
                  mrmodel = ~ glm(link = 'logit', formula = ~ size), #  distance),
                  data = picnic.data1.swapObs,
                  meta.data = list(width = 3))


detection.tables.swap <- det.tables(fi.mr.dist.swap)
# Print these detection tables
print(detection.tables.swap)
plot(fi.mr.dist.swap)

```


2026.06.01 Michelle DePrenger-Levin additions from Hamilton et al. 2017   
inverse logit
```{r}
## Then repeat by swapping EW and NS? As trial - different form fully independent
## x == distance
## s == group size

# Probability of EW detection 
y_EW <- exp(alpha1 + beta1*x + gamma1*(x*x) + delta1*s + epsilon1*x*s) / 
          (exp(alpha1 + beta1*x + gamma1*(x*x) + delta1*s + epsilon1*x*s)+1)

# Probability of detection by NS conditional on non-detection of EW 
y_EW_notNS <-  exp(alpha2 + beta2*x + gamma2*(x*x) + delta2*s + epsilon2*x*s) / 
          (exp(alpha2 + beta2*x + gamma2*(x*x) + delta2*s + epsilon2*x*s)+1)

# Probability of detection by NS conditional on detection by EW (duplicates)
y_EW_NS <-  exp(alpha3 + beta3*x + gamma3*(x*x) + delta3*s + epsilon3*x*s) / 
          (exp(alpha3 + beta3*x + gamma3*(x*x) + delta3*s + epsilon3*x*s)+1)


```

