---
title: "MRDS for Crossed SCGL"
author: "Michelle DePrenger-Levin"
date: "2026-01-25"
output: html_document
---

```{r message=FALSE, warning=FALSE, include=FALSE}

rm(list =ls())

# tidyverse contains both 'dyplr' and 'ggplot2'
library(tidyverse)
library(dplyr)
library(janitor)
library(ggrepel)


library(devtools)
library(terra)
library(raster)
library(ggplot2)

## MRDS
library(unmarked)
library(MuMIn)
library(Distance)  
library(dssd)

setwd("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScleroDistanceSampling")
scgl_obs <- read_csv("Data/distanceSampling_1.csv")
scgl_trans_length <- read_csv("Data/transect_2.csv")

# scgl_obs <- read_csv("data_raw/distanceSampling_1.csv")
# scgl_trans_length <- read_csv("data_raw/transect_2.csv")

scgl_obs <- clean_names(scgl_obs)
```


This function will be used later to find duplicate plants    
Update to same group size 


```{r test function}

# `site` = a dataframe with relative (x,y) coords as seperate columns `x` & `y`, transect directions as `NS` & `EW`, and unique `plant_id' for each observation.
# `b` = an allowance in meters(m) within which two observations (1 NS & 1 EW) will be considered the same observation  
# Also account for differences in group size - for now, take larger
crossed <- function(site, b){
  
 temp_site <-  site %>% 
    rowwise() %>% 
    mutate(coord = list(c(x,y)))
  
#chart all possible combinations of NS coordinates and EW coordinates
 grid <-  expand.grid(EW = temp_site[which( site$transect_orientation == "EW"), "coord"]$coord,
                      NS = temp_site[which( site$transect_orientation == "NS"), "coord"]$coord) %>% 
    rowwise() %>% # to look for the differences between EW and NS points by row
   #make a column of distances between points
    mutate( dist = sqrt(sum((EW - NS)^2))) %>% 
   #filter the combinations to only the ones within the supplied fuzziness `b`
    filter(dist <= b)

#appending to contain columns with EW plant_id's for each combination ## What if we reverse the order: NS first, then EW
 grid <- left_join(grid, temp_site[, c("coord", "plant_id")],
                   by = join_by(EW == coord)) %>%
   mutate(EW_ID = plant_id, plant_id = NULL) %>%
  #appending to contain columns with NS plant_id's for each combination
   left_join(temp_site[, c("coord", "plant_id")], 
                   by = join_by(NS == coord)) %>% 
   mutate(NS_ID = plant_id, plant_id = NULL) 

#checking for Id's with multiple matches.(Should be only one distinct NS observation for each distinct EW observation). Keep only the match with the least distance between points
if(!(all(count(grid, NS_ID)$n >1) | all(count(grid, EW_ID)$n >1)) ){
  grid <- filter(group_by(grid, NS_ID), dist == min(dist))
  grid <- filter(group_by(grid, EW_ID), dist == min(dist))
  }
 
 #Within the original dataframe, change corresponding EW ID's to their NS counterparts
 # temp_site <- left_join(temp_site, grid[, c("EW_ID", "NS_ID")], by = join_by(plant_id == EW_ID)) %>% 
 #   mutate( plant_id = ifelse( !is.na(NS_ID), NS_ID, plant_id), NS_ID = NULL, coord = NULL) %>% 
 #   ungroup()
 
 ## More if swapped!
 temp_site <- left_join(temp_site, grid[, c("EW_ID", "NS_ID")], by = join_by(plant_id == NS_ID)) %>% 
   mutate( plant_id = ifelse( !is.na(EW_ID), EW_ID, plant_id), EW_ID = NULL, coord = NULL) %>% 
   ungroup()
 
 temp_site
 
}

# temp_site %>%
#   group_by(plant_id, transect_orientation) %>%
#   mutate(n = n()) %>%
#   arrange(plant_id) %>%
#   filter(n >1)

# site %>%
#   select(distance_to_group, transect_meter_mark, x, y)

```


Start here
```{r}
scgl <- scgl_obs[!(is.na(scgl_obs$transect_meter_mark|scgl_obs$distance_to_group)),] %>% #filtering NA's and outliers. 
        filter(transect_meter_mark < 110, distance_to_group < 10) %>% 
        mutate( object_id = as.factor(object_id),
                site_name = as.factor(site_name),
                transect_number = as.factor(transect_number))

            
scgl1 <- scgl %>% 
   mutate(
     # EW transect meter marks need to be * -1 because they were recorded as if in the(+,+) quadrant, but the mapping we will do is based on (-,+) quadrant. 
    transect_meter_mark = case_when( 
      transect_orientation == "EW" ~ transect_meter_mark * -1, 
      TRUE ~ transect_meter_mark),
    # making the 'direction to group' indicate if the 'distance to group' needs to be left/right of transect.  
    distance_to_group = case_when(
      direction_to_group == "-" ~ distance_to_group * -1, 
      TRUE ~ distance_to_group ) 
          )
           

# setting up the (X,Y) columns to then transform based on plot/shapefile analysis
scgl1 <- scgl1 %>% 
  mutate( x = case_when(transect_orientation == "NS" ~ distance_to_group, transect_orientation == "EW" ~ transect_meter_mark), 
          y = case_when(transect_orientation == "EW" ~ distance_to_group, transect_orientation == "NS" ~ transect_meter_mark)) 

scgl1 <- left_join(scgl1, scgl_trans_length[, c("Transect Length", "GlobalID")], by = join_by(parent_global_id == GlobalID))                
                
#filtering to sites
fram <- scgl1 %>% 
  filter(site_name == "12")

picnic <- scgl1 %>% 
  filter(site_name == "1"| site_name == "3")

bridgeport <- scgl1 %>% 
  filter(site_name == "10")

oilpad <- scgl1 %>% 
  filter(site_name == "11")


write.csv(picnic, file = "C:/Users/deprengm/ScleroDistanceSampling/Data/picnicCrossed.csv")
```




Conventional distance sampling for NS and EW, treat them like repeat samples  
```{r}

## Transects from the grid running NS or grid running EW
## Strata are the two perpendicular grids
## distance: perpendicular distance from the transect line
## size: number of individuals within a group 
## status: 1 if any are reproductive, 0 if all are vegetative
## length of that particular transect

CDS <- picnic %>%
  dplyr::select(transect_number, transect_orientation, distance_to_group, 
         number_seedlings_in_group, number_vegetative_in_group, number_reproductive_in_group,
         `Transect Length`) %>%
  dplyr::rename(length = 'Transect Length') %>% 
  dplyr::mutate(distance = abs(distance_to_group)) %>%
  rowwise() %>%
  dplyr::mutate(size = sum(c(number_seedlings_in_group, number_vegetative_in_group, number_reproductive_in_group), na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::mutate(strata = transect_orientation) %>%
  dplyr::mutate(transect = paste0(strata, transect_number)) %>%
  ## Reproductive Status yes=1, no=0
  dplyr::mutate(status = case_when(number_reproductive_in_group > 0 ~ 1,
                            TRUE ~ 0)) %>%
  dplyr::mutate(status = as.factor(status)) %>%
  dplyr::mutate(length = case_when(transect_number == 1 & strata == "EW" ~ 70,
                            transect_number == 2 & strata == "NS" ~ 66,
                            TRUE ~ length)) %>%
  filter(size > 0) %>%
  dplyr::select(transect, strata, distance, size, status, length) # %>%
  # mutate(transect = as.factor(transect),
  #        strata = as.factor(strata))
  

lengths <- CDS %>%
  distinct(transect, strata, length) %>%
  arrange(strata, transect)

lengths %>%
  group_by(strata) %>%
  summarise(L = sum(length))
# 203 meters for EW
# 156 meters for NS
```

<https://cran.r-project.org/web/packages/unmarked/vignettes/distsamp.html>  

There are more EW than NS 
```{r}
library(patchwork)
## Do other covariates impact detections?
CDS %>%
  ggplot( aes(size, distance, color = strata))+
    geom_point(pch = 16) +
    geom_smooth() +
    scale_color_manual(values = c("#1F70B7", "#E69F00"))  # palette.colors(8)

CDS %>% 
  mutate(DistClass = ifelse(distance >=1.5, "far", "near")) %>%
  ggplot( aes(size, fill = strata))+
    geom_histogram(binwidth = 0.5, position = "dodge") +
    scale_fill_manual(values = c("#1F70B7", "#E69F00")) +
    theme_bw() +
    scale_x_continuous(breaks = 1:10) +
    facet_wrap(~ DistClass)

CDS %>% 
  # Add cut points
  mutate(distClass = as.numeric(cut(distance, seq(0, 3, 0.25)))) %>%
  ggplot( aes(distance, size, color = status))+
    geom_point() +
    scale_color_manual(values = c("#1F6D58", "#CC79A7")) +
    theme_bw() +
    facet_wrap(~strata) +
    geom_smooth(method = "lm", se = FALSE)

CDS %>% 
  # Add cut points
  mutate(distClass = cut(distance, seq(0, 3, 0.25))) %>%
  ggplot( aes(distance, fill = status))+
    geom_histogram(bins = 20, position = "dodge") +
    scale_fill_manual(values = c("#1F6D58", "#CC79A7")) +
    theme_bw() +
    facet_wrap(~strata) 

CDS %>% 
  # Add cut points
  mutate(distClass = cut(distance, seq(0, 3, 0.25))) %>%
  ggplot( aes(distance, fill = strata))+
    geom_histogram(bins = 20, position = "dodge") +
    scale_fill_manual("Distance\nClass",values = c("#A78623", "#4A8CC5")) +
    theme_bw()  

CDS %>% 
  mutate(DistClass = ifelse(distance >=1.5, "far", "near")) %>%
  ggplot( aes(size, fill = DistClass))+
    geom_histogram(binwidth = 0.5, position = "dodge") +
    scale_fill_manual("Distance\nClass",values = c("#A78623", "#4A8CC5")) +
    theme_bw() +
    scale_x_continuous(breaks = 1:10) +
    facet_wrap(~ strata)

## Shape is similar (pooling robust) but the scale is different
g1 <- CDS %>%
  ggplot( aes(distance, fill = strata)) +
    geom_histogram() +
    theme_bw() +
    scale_fill_manual(values = c("#1F70B7", "#004D40"), guide = "none") +
    xlab("Distance (m)")

g2 <- CDS %>%
  ggplot( aes(status, distance, fill = strata)) +
    geom_boxplot(position = position_dodge(width = 1)) +
    geom_point(position = position_jitterdodge(dodge.width = 1, jitter.width = 0.5, jitter.height = 0), pch = 16) +
    scale_x_discrete("", labels = c("0" = "Vegetative", "1" = "Reproductive")) +
    scale_fill_manual("Orientation",values = c("#1F70B7", "#004D40")) +
    theme_bw() +
    ylab("Distance (m)")


## How many hits among the EW vs. NS
CDS %>%
  group_by(strata) %>%
  summarise(n=n())

# Switch the histogram with the probability density differences between NS and EW
ggsave(filename = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Sclerocactus-glaucus/Sclerocactus-glaucus_Manuscripts/DistanceSampling_Images/CrossedHistDistXRepro.jpg",
g1 + g2 + plot_layout(widths = c(0.75, 1)),
width=150, height=70,units='mm', dpi=300)

```

unmarked   
Assuming all are individuals, estimting the number of clusters. 
```{r}
library(HDInterval)
CDS %>%
  group_by(strata) %>%
  summarise(clustersize = mean(size), ## 1.56
            lcl = hdi(size)[1],       ## 1 
            hcl = hdi(size)[2])       ## 4

CDS %>%
  ggplot(  aes(size, fill = strata, alpha = 0.5)) +
    geom_density() +
    geom_vline(xintercept = 1.56, linetype = "dotted", lwd = 1) +
    geom_vline(xintercept = 4) + 
    geom_vline(xintercept = 1) +
    theme_bw() +
    scale_fill_manual("Strata", values = c("#1F6D58", "#CC79A7")) +
    scale_alpha(guide = "none")

yDat <- unmarked::formatDistData(as.data.frame(CDS), distCol = "distance",
                       transectNameCol = "transect", dist.breaks = c(seq(0,3, 0.25)))

covs <- data.frame(Strata = unlist(lapply(row.names(yDat), function(x) gsub("[[:digit:]]","",x))))
covs$Strata <- as.factor(covs$Strata)
covs$transect <- as.factor(row.names(yDat))

tlength <- CDS %>%
  dplyr::distinct(transect, length)

## Need to average, not sum two strata
Nhat <- covs %>%
  left_join(tlength) %>%
  mutate(NS = ifelse(Strata == "EW", 1, 0),
         EW = ifelse(Strata == "NS", 1, 0))

umf <- unmarkedFrameDS(y = as.matrix(yDat), 
                       siteCovs = covs, 
                       survey = "line",
                       dist.breaks =  c(seq(0,3, 0.25)),
                       tlength = tlength$length,
                       unitsIn = "m")

summary(umf)

hist(umf, xlab = "distance (m)")

hn_Null <- distsamp(~1 ~1, umf, keyfun="halfnorm", output="density",unitsOut="kmsq")
hn_Strata <- distsamp(~1 ~Strata, umf, keyfun="halfnorm", output="density",unitsOut="kmsq")
hn_StrataStrata <- distsamp(~Strata ~Strata, umf, keyfun="halfnorm", output="density",unitsOut="kmsq")
# If parameters are at boundary - use half normal instead
# haz_Null <- distsamp(~1 ~1, umf, keyfun="hazard", output="density",unitsOut="kmsq", starts = c(0.5,0.5))
# haz_Strata <- distsamp(~1 ~Strata, umf, keyfun="hazard", output="density",unitsOut="kmsq")
# haz_Strata <- distsamp(~Strata ~Strata, umf, keyfun="hazard", output="density",unitsOut="kmsq")

names(hn_Null)
MuMIn::AICc(hn_Null, hn_Strata, hn_StrataStrata)

require(AICcmodavg)
distsamp.list <- list(hn_Null, hn_Strata, hn_StrataStrata)
distsamp.names <- as.character(unlist(lapply(distsamp.list, formula)))
(distsamp.results <- aictab(distsamp.list, modnames = distsamp.names))

# evidence ratio
for(i in 2:length(distsamp.list)){
  print(exp(0.5*distsamp.results$Delta_AICc[i]))
}


## Predict to get estimates and 95% confidence intervals from the model on the original scale
predict(hn_Null, type = "state")[1,]/1000000
predict(hn_Strata, type = "state")/1000000

# shape
predict(hn_Null, type = "det")  # Same
predict(hn_Strata, type = "det")
predict(hn_StrataStrata, type = "det")


# scale - only for hazard rate
# predict(hn_Null, type = "scale")[1,]


nd <- data.frame(Strata = "NS")

predict(hn_Strata, type = "state", newdata = nd, appendData = TRUE)


## Number of individuals in the sampled plots
```

Uncertainty of derived parameters, use parametric bootstrap approach. Define function, then apply function to numerous models fit to data simulated from original model     

https://distancesampling.org/Distance/articles/web-only/differences/differences.html   Bootstrap
https://cran.r-project.org/web/packages/unmarked/vignettes/distsamp.html     Code below from here
```{r}
## The population in the covered area of transects
getN.hat <- function(fit) {
    d <- predict(fit, type="state")$Predicted
    a.NS <- d * ((Nhat$length * 6 * Nhat$NS)/ 1000000)
    a.EW <- d * ((Nhat$length * 6 * Nhat$EW)/ 1000000)
    N.hat <- c(N.hat = sum( (a.NS + a.EW)/2 ))
    return(N.hat)
    }
pb <- parboot(hn_Strata, statistic=getN.hat, nsim=999)
pb

## From projected population 
getN.hatTot <- function(fit) {
    dNS <- predict(fit, type="state")$Predicted[which(Nhat$NS == 1)[1]]
    dEW <- predict(fit, type="state")$Predicted[which(Nhat$EW == 1)[1]]
    a.NS <- dNS * (7807/ 1000000)
    a.EW <- dEW * (7807/ 1000000)
    N.hat <- c(N.hat = sum( (a.NS + a.EW)/2 ))
    return(N.hat)
    }
pbTot <- parboot(hn_Strata, statistic=getN.hatTot, nsim=999)
pbTot

# Density
getD.hatTot <- function(fit) {
    dNS <- predict(fit, type="state")$Predicted[which(Nhat$NS == 1)[1]]
    dEW <- predict(fit, type="state")$Predicted[which(Nhat$EW == 1)[1]]
    a.NS <- dNS / 1000000
    a.EW <- dEW / 1000000
    N.hat <- c(N.hat = sum( (a.NS + a.EW)/2 ))
    return(N.hat)
    }
pbDen <- parboot(hn_Strata, statistic=getD.hatTot, nsim=999)
pbDen
```


### Plotting individuals  
Reference Buckland et al. 2009 to talk about difference in detection function vs. pooled
```{r}

strataCon <- data.frame(Strata = factor(c("EW","NS")))

# Estimate density and sigma for each row
Elambda <- predict(hn_Strata, type = "state", newdata = strataCon, appendData = TRUE)
Esigma <- predict(hn_Strata, type = "det", newdata = strataCon, appendData = TRUE)

par(mfrow=c(1, 1))
plot(function(x) gxhn(x, sigma = 1.3456), 0, 3, xlab = "Distance (m)",
     ylab = "Detection Prob", cex.lab = 0.7,
     cex.axis = 0.7, las = 1)
hist(hn_Strata, xlab = "Distance (m)", ylab = "Probability density", main = "")
# dev.off()

## Detection function begins with "g" and density function begins with "d"
## For hn_StrataStrata   c("#1F70B7", "#004D40")
jpeg("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Sclerocactus-glaucus/Sclerocactus-glaucus_Manuscripts/2026_Ecological-Modelling/Images/EWvsNSPooledDetectionfxn.jpg",
     width = 210, height = 100, units = "mm",res=300)
par(mfrow=c(1, 2))
# plot(function(x) gxhn(x, sigma = 1.3456), 0, 3, xlab = "Distance (m)",
#      ylab = "Detection Prob", cex.lab = 0.7,
#      cex.axis = 0.7, las = 1)
hist(hn_Strata, xlab = "Distance (m)", ylab = "Probability density", main = "") #, ylim = c(0,1))
plot(function(x) dxhn(x, sigma = 1.3909), 0,3,add=TRUE,  lwd = 2,col = "#1E88E5", lty = "dashed")  ## gxhn is the detection function and g(0) = 1
plot(function(x) dxhn(x, sigma = 1.26471), 0,3, lwd = 2, add=TRUE, col = "#D81B60", lty = "dotted")
legend('topright', c("EW","Pooled","NS"),
       col = c("#1E88E5","black", "#D81B60"), lty=c("dashed","solid","dotted"), lwd = 2,cex=1)

## Detection function
plot(function(x) gxhn(x, sigma = 1.345641), 0,3, lwd = 2, col = "black", xlab = "Distance (m)", ylab = "Detection function (half-normal)")
plot(function(x) gxhn(x, sigma = 1.3909), 0,3,add=TRUE,  lwd = 2,col = "#1E88E5", lty = "dashed")  ## gxhn is the detection function and g(0) = 1
plot(function(x) gxhn(x, sigma = 1.26471), 0,3, lwd = 2, add=TRUE, col = "#D81B60", lty = "dotted")
legend('topright', c("EW","Pooled","NS"),
       col = c("#1E88E5","black", "#D81B60"),lty=c("dashed","solid","dotted"), lwd = 2,cex=1)
dev.off()


```

https://examples.distancesampling.org/DSsim-truncation/DSsim-examples.html  




Involves transforming the coordinates to match the plot design.

```{r picnic transform and plot}

#describing the plot design
pic_lines <- tibble( x = c(-20, -30, -40, -20, -10, -10, 0), 
                     y = c(20, 10, 0, 20, 30, 40, 50), 
                     xend = c(-20, -30, -40, -90, -70, -70, -50),
                     yend = c(60, 76, 53, 20, 30, 40, 50),
                     label = c("1NS", "2NS", "3NS", "1EW", "2EW", "3EW", "4EW"))
                
 picnic_1 <-  picnic %>% 
   mutate( 
    x = case_when(transect_orientation == "NS" & transect_number == "1" ~ x - 20, 
                  transect_orientation == "NS" & transect_number == "2" ~ x - 30,
                  transect_orientation == "NS" & transect_number == "3" ~ x - 40,
                  transect_orientation == "EW" & transect_number == "1" ~ x - 20,
                  transect_orientation == "EW" & transect_number == "2" ~ x - 10,
                  transect_orientation == "EW" & transect_number == "3" ~ x - 10,
                  transect_orientation == "EW" & transect_number == "4" ~ x
                  ),
    y = case_when(transect_orientation == "NS" & transect_number == "1" ~ y + 20, 
                  transect_orientation == "NS" & transect_number == "2" ~ y + 10,
                  transect_orientation == "NS" & transect_number == "3" ~ y ,
                  transect_orientation == "EW" & transect_number == "1" ~ y + 20,
                  transect_orientation == "EW" & transect_number == "2" ~ y + 30,
                  transect_orientation == "EW" & transect_number == "3" ~ y + 40,
                  transect_orientation == "EW" & transect_number == "4" ~ y + 50,
                  )
   )
 
pic_plot <-  picnic_1 %>% 
  ggplot( mapping = aes(x = x, y = y)) +
  geom_point(aes(color = transect_orientation, shape = transect_orientation)) +
  geom_segment(data = pic_lines, aes(x = x, y = y, xend = xend, yend = yend)) +
  theme_bw() +
  scale_color_discrete("Strata", type = c("#D81B60","#1E88E5"))+
  scale_shape_manual("Strata", values = c(4,1)) +
  coord_equal(ratio = 1)

ggsave(filename = "PicnicPlot.jpg", pic_plot ,
width=150, height=125,units='mm', dpi=300) 


```




### Marking Overlap

```{r}

#apply unique ID's
picnic_id <- dplyr::mutate(picnic_1, plant_id = c(1:nrow(picnic_1)))

#filter to crossing zones
## Increase to + or - 4 from 3 (that was real) to account for ones that are off due to non right angles?
cD <- 3 ## cross distance
picnic_id <- dplyr::mutate(picnic_id, 
                    zoned = case_when( 
  c(rowSums(mapply( FUN = function(a, b){between(a, b-cD, b+cD)}, picnic_id[, "y"], c(20, 30, 40, 50))) +
      rowSums(mapply( FUN = function(a, b){between(a, b-cD, b+cD)}, picnic_id[, "x"], c(-20, -30, -40)))) == 2 ~ 1,
  TRUE ~ 0)) 


#find duplicates within those zones; increase distance of fuzziness
picnic_duped <- crossed(picnic_id, 1)

## Duplicate observations; there are 10, some don't match in group size or composition
picnic_duped %>%
  group_by(plant_id, zoned) %>%
  mutate(n = n()) %>%
  # filter(duplicated(plant_id)) %>%
  dplyr::select(plant_id, transect_orientation, number_seedlings_in_group:number_reproductive_in_group, n) %>%
  filter(n >1) %>%
  arrange(plant_id)

picnic_duped %>%
  group_by(zoned) %>%
  distinct(plant_id) %>%
  summarise(n = n())

pic_plot1 <- picnic_duped %>% 
  filter(zoned ==1) %>% 
  ggplot( mapping = aes(x, y)) +
  geom_point(aes(color = transect_orientation), shape = 1) +
  geom_point(data = filter(picnic_duped, duplicated(plant_id)), color = "black") +
  geom_segment(data = pic_lines, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_text_repel(data = pic_lines,
            aes( label = label),
            size = 2,
            ) +
  coord_equal(ratio = 1) +
  scale_color_manual("", values =  c("#D81B60","#1E88E5"))  # +
   # geom_text_repel( data = filter(picnic_duped, duplicated(plant_id)), aes( label = plant_id), size = 3, min.segment.length = 0.1)  # label the overlapping observations


pic_plot1

ggsave(filename = "PicnicPlotCrossed.jpg", pic_plot1 ,
width=150, height=125,units='mm', dpi=300) 


```

Lincoln-Petersen Index of population size (this doesn't get detection)
```{r}
picnic_duped %>%
  filter(zoned == 1) %>%
  summarise(n = n())

# Marked individuals released 
# M <- 
  
library(RMark)

## Independent double observer transect survey; Huggins closed capture, Observer 1 was SL (spotlight), TI (thermal imager)
# Group similar to transect? or grid; Cluster assigns deer to a specific observation cluster
# Collier et al 2012
data(deer)
  
```

Mark-recapture  
```{r}
x <- picnic_duped %>%
  filter(zoned == 1) %>%
  ## Detected
  mutate(dh = 1) %>%
  # group_by(plant_id, transect_orientation) %>%
  complete(transect_orientation, plant_id,
                fill = list(dh = 0)) %>%
  dplyr::select(dh, plant_id, transect_orientation) %>%
  pivot_wider(id_cols = plant_id, values_from = dh, names_from = transect_orientation) %>%
  unite(ch, EW:NS, sep="") # 

x2 <- picnic_duped %>%
  filter(zoned == 1) %>%
  ## Detected
  mutate(dh = 1) %>%
  # group_by(plant_id, transect_orientation) %>%
  complete(transect_orientation, plant_id,
                fill = list(dh = 0)) %>%
  dplyr::select(dh, plant_id, transect_orientation) %>%
  pivot_wider(id_cols = plant_id, values_from = dh, names_from = transect_orientation) %>%
  unite(ch, NS:EW, sep="") # 

x %>%
  mutate(ch = paste0(ch,";")) %>%
  write.table(file = "Data/Closed.inp",
              sep = " ", col.names = FALSE, row.names = FALSE)
  



## Convert for mark
scglClosed.inp <- convert.inp("Data/Closed.inp",
                              covariates = NULL,
                              use.comments = FALSE)

M <- nrow(x) ## 81 unique observations
```

Huggins Closed so I could have individual covariates but don't need them    
Closed: constant p, constant c, f0 = frequency (number) never seen (need to add unique number captured to estimate population)   
Know what the number of unique observations is, make sure N hasn't collapsed to the value of unique observations 
```{r}
## EW then NS
Cproc <- process.data(x, model = "Closed")
Cddl <- make.design.data(Cproc)

# constant p and c
closed <- mark(data = Cproc, Cddl, model.parameters = list(p = list(formula = ~1),
                                                           c = list(formula = ~1)))

closedshared <- mark(data = Cproc, Cddl, model.parameters = list(p = list(formula=~1,share=TRUE)))
## NS then EW
Cproc2 <- process.data(x2, model = "Closed")
Cddl2 <- make.design.data(Cproc2)

# constant p and c
closed2 <- mark(data = Cproc2, Cddl2, model.parameters = list(p = list(formula = ~1),
                                                              c = list(formula = ~1)))


closedshared2 <- mark(data = Cproc2, Cddl2, model.parameters = list(p = list(formula=~1,share=TRUE)))


## Huggins 
## EW then NS
CprocHug <- process.data(x, model = "Huggins")
CddlHug <- make.design.data(CprocHug)

# constant p and c
Huggins <- mark(data = CprocHug, CddlHug, model.parameters = list(p = list(formula = ~1),
                                                           c = list(formula = ~1)))
## NS then EW
Cproc2Hug <- process.data(x2, model = "Huggins")
Cddl2Hug <- make.design.data(Cproc2Hug)

# constant p and c
Huggins2 <- mark(data = Cproc2Hug, Cddl2Hug, model.parameters = list(p = list(formula = ~1),
                                                              c = list(formula = ~1)))


get.real(Huggins, "p", se = TRUE)
get.real(closed, "p", se = TRUE)
get.real(closed, "f0", se = TRUE)
get.real(closedshared, "p", se = TRUE)

get.real(Huggins2, "p", se = TRUE)
get.real(closed2, "p", se = TRUE)
get.real(closedshared2, "p", se = TRUE)
get.real(closedshared, "f0", se = TRUE)


## 121.98 (56.5 - 263.3) as the 95% CI of individuals never seen

```



```{r}
# half normal model with key function = 
exp(-animals$perd^2/(2 * animals$sigma^2)). 

# For a hazard rate model, I replaced the key function with: 
1-exp(-(animals$perd/animals$sigma) ^ -b) 
# where b is a shape parameter given in summary(mod)$ddf, but left the rest of the code as is. Is this correct?

I figured out a 'clever' way of extracting the model and shape parameter coefficients for my hazard rate model: 
a <- coef(mod$ddf)[1]$scale$estimate
b <- coef(mod$ddf)[2]$exponent$estimate
```


Follow example and make data like mrds 
Independent observer point independence model     
io.fi: the mark-recapture component of the modelling  

```{r}
library(mrds)

picnic.data <- picnic_duped %>%
  filter(zoned == 1) %>%
  mutate(detected = 1) %>%
  mutate(distance = abs(distance_to_group)) %>%
  rowwise() %>%
  mutate(size = sum(c(number_seedlings_in_group, number_vegetative_in_group, number_reproductive_in_group), na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::rename("observer" = "transect_orientation") %>%
  dplyr::select(plant_id, observer, detected, ## detected by fill all plant_id for each trans_ori
         distance, size, transect_meter_mark, transect_number) %>%
  complete(nesting(plant_id,size),observer, fill = list(detected = 0)) %>% ## Keep only size and plantID that are in the data
  relocate(size, .after = distance) %>%
  ## When grouped, one might be missed, that's the row where distance missing
  group_by(plant_id) %>%
  ## each set of two observations
  ## need to update the group size as well. 
  mutate(distance2 = case_when(detected == 0 & max(abs(transect_meter_mark), na.rm = TRUE) < 15 ~ abs(10-max(abs(transect_meter_mark), na.rm = TRUE)),
                        detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 15 & max(abs(transect_meter_mark), na.rm = TRUE) < 25 ~ 
                                                                                                 abs(20-max(abs(transect_meter_mark), na.rm = TRUE)),
                        detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 25 & max(abs(transect_meter_mark), na.rm = TRUE) < 35 ~ 
                                                                                                 abs(30-max(abs(transect_meter_mark), na.rm = TRUE)),
                        detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 35 & max(abs(transect_meter_mark), na.rm = TRUE) < 45 ~ 
                                                                                                 abs(40-max(abs(transect_meter_mark), na.rm = TRUE)),
                        detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 45 & max(abs(transect_meter_mark), na.rm = TRUE) < 55 ~ 
                                                                                                 abs(50-max(abs(transect_meter_mark), na.rm = TRUE)),
                        detected == 0  & max(abs(transect_meter_mark), na.rm = TRUE) > 55 & max(abs(transect_meter_mark), na.rm = TRUE) < 65 ~ 
                                                                                                 abs(60-max(abs(transect_meter_mark), na.rm = TRUE)),
      TRUE ~ max(distance, na.rm = TRUE)
                              )) %>%
  # ungroup() %>%
  dplyr::select(plant_id:detected,size,distance2) %>%
  rename("distance" = "distance2")

picnic_duped %>%
  dplyr::select(plant_id, number_seedlings_in_group:number_reproductive_in_group,transect_orientation) %>%
  dplyr::arrange(plant_id)


picnic_duped %>%
  filter(plant_id == 6) %>%
  dplyr::select(number_seedlings_in_group:number_reproductive_in_group)


## Shouldn't be two observations of same plant id in same line
picnic.data %>%
  group_by(plant_id, observer) %>%
  summarise(n = n()) %>%
  filter(n > 1)

# When the group size differs, then 
## One is one plant, one are two plants; discrepancies in group size. 
picnic.data %>%
  filter(plant_id == 41)
# Two different groups likely
picnic_duped %>%
  filter(plant_id == 41) %>%
  dplyr::select(number_seedlings_in_group:number_reproductive_in_group)
## This one seems like should be two different plants. Unlikely to confuse 1 reproductive with 3 vegetative
picnic_duped %>%
  filter(plant_id == 136) %>%
  dplyr::select(number_seedlings_in_group:number_reproductive_in_group)
picnic_duped %>%
  filter(plant_id == 141) %>%
  dplyr::select(number_seedlings_in_group:number_reproductive_in_group)
# This one is likely the same plant, both have one reproductive, one has one seedling. 
picnic_duped %>%
  filter(plant_id == 188) %>%
  dplyr::select(number_seedlings_in_group:number_reproductive_in_group)

## Keep only one repeated plant_id per EW and NS pair.
picnic.data1 <- picnic.data %>%
  # group_by(plant_id, observer) %>%
  # distinct(plant_id, .keep_all = TRUE) %>%
  # arrange(-detected) %>%
  # mutate(distance = case_when(detected == 0 ~ NA,
  #                             TRUE ~ distance)) %>%
  # ungroup() %>%
  # group_by(plant_id) %>%
  # fill(distance) %>%
  # ungroup() %>%
  rename("object" = "plant_id") %>%
  arrange(object, observer) %>%
  mutate(observer = as.factor(as.numeric(as.factor(observer))))

picnic.data1 %>%
  group_by(object) %>%
  summarise(n = n()) %>%
  filter(n > 1)
  
picnic.region <- data.frame(Region.Label = "Picnic", 
                            Area = 7802)

picnic.samples <- picnic %>%
  distinct(transect_number, transect_orientation, `Transect Length`) %>%
  rowwise() %>%
  mutate(Sample.Label = paste0(transect_number,transect_orientation)) %>%
  ungroup() %>%
  mutate(Region.Label = "Picnic") %>%
  rename("Effort" = "Transect Length") %>%
  select(Sample.Label, Region.Label, Effort) %>%
  mutate(Effort = case_when(Sample.Label == "1EW" ~ 70,
                            TRUE ~ Effort)) %>%
  filter(!is.na(Effort)) %>%
  arrange(Sample.Label)

picnic.obs <- picnic_duped %>%
  filter(zoned == 1) %>%
  rowwise() %>%
  mutate(Sample.Label = paste0(transect_number,transect_orientation)) %>%
  ungroup() %>%
  distinct(plant_id, Sample.Label) %>%
  mutate(Region.Label = "Picnic")

```


<https://distancesampling.org/mrds/articles/mrds-golftees.html>     
Not appropriate for this data because the distances differ so not an estimate of how detection differs by distance
```{r}

## Fully independent   
# Should break into distance categories which isn't the same for our data anyway, might be close for 
fi.mr.dist <- ddf(method = "trial.fi", 
                  mrmodel = ~ glm(link = 'logit', formula = ~ distance),
                  data = picnic.data1,
                  meta.data = list(width = 3))



```



```{r}

ds.data <- picnic %>%
  select(number_seedlings_in_group:number_reproductive_in_group, 
         transect_number, transect_orientation,distance_to_group) %>%
  rowwise() %>%
  mutate(size = sum(c(number_seedlings_in_group, 
                      number_vegetative_in_group, 
                      number_reproductive_in_group), na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(repro = case_when(number_reproductive_in_group > 0 ~ 1, 
                           TRUE ~ 0)) %>%
  select(transect_number:repro) %>%
  mutate(distance = abs(distance_to_group)) %>%
  rename("strata" = "transect_orientation") %>%
  mutate(strata = as.factor(strata)) %>%
  select(-distance_to_group)
  


```

